---
title: "Anchovy embrios Aegean barrier Mesh sensitivity "
author: "Eleni Servou"
date: today
format:
  html:
    toc: true
    toc-depth: 4
    toc-float: true
    code-fold: true         
    code-summary: "Hide/Show "
    code-tools:
      toggle: true                
---

## Load libraries and prepare data
```{r}
# Load packages
library(tidyverse)
library(dplyr)
library(ggplot2)
library(mapdata)
library(rnaturalearth)
library(rnaturalearthdata) 
library(sf)
library(sp)
library(sdmTMB)
library(sdmTMBextra)
library(INLA)
library(DHARMa)
library(gstat)
library(readxl)
library(gt)
library(visreg)
library(ggeffects)
library(cli)
```

```{r}
#Data set (Eggs and larvae at age) 
Str1_19<- read_xlsx("data.xlsx", sheet= "2019 east") ;Str1_19$Survey <- "Str1_19"
Str2_19<- read_xlsx("data.xlsx", sheet= "2019 west") ;Str2_19$Survey <- "Str2_19"
Str1_19$Survey <- "Str1_19"
Str2_19$Survey <- "Str2_19"
```

```{r}
#Data for maps
#Coastline
coast <- st_read("Greece_geology/Greece/land_poly_GR.gpkg",quiet = TRUE) %>%  st_transform(crs = "WGS84") 
coast_km<-  st_transform(coast, crs = 32635);  
st_geometry(coast_km) <- st_geometry(coast_km) / 1000  
```

## Estimate stations weights 
```{r}

 Str1_19_w<- Str1_19 %>%
   distinct(Station, .keep_all = TRUE) %>%  mutate(total_km = sum(km_2, na.rm = TRUE)) %>%
   mutate(w_sn=km_2/mean(km_2,na.rm=TRUE))
   
 Str1_19 <- Str1_19 %>%
   left_join(Str1_19_w %>% select(Station,w_sn), by = "Station")  %>%
  add_utm_columns(ll_names = c("Lon","Lat"), ll_crs = "WGS84",
                  utm_crs = 32635, utm_names = c("X","Y"), units = "km")
 
Str2_19_w<- Str2_19 %>%
   distinct(Station, .keep_all = TRUE) %>%  mutate(total_km = sum(km_2, na.rm = TRUE)) %>%
   mutate(w_sn=km_2/mean(km_2,na.rm=TRUE))
   
Str2_19 <- Str2_19 %>%
   left_join(Str2_19_w %>% select(Station, w_sn), by = "Station")  %>%
  add_utm_columns(ll_names = c("Lon","Lat"), ll_crs = "WGS84",
                  utm_crs = 32635, utm_names = c("X","Y"), units = "km")
```

#Load and plot polygons for each stratum and year
```{r}
PA_2019_Str1 <- st_read("voronois/str1_sd_2019_dis.shp")  %>%  st_transform(crs = "WGS84")
PA_2019_Str2 <- st_read("voronois/str2_sd_2019_dis.shp")  %>%  st_transform(crs = "WGS84")

PA_2019_Str1 <- PA_2019_Str1 %>%  st_transform( crs = 32635) 
st_geometry(PA_2019_Str1) <- st_geometry(PA_2019_Str1) / 1000 

PA_2019_Str2 <- PA_2019_Str2 %>%  st_transform( crs = 32635) 
st_geometry(PA_2019_Str2) <- st_geometry(PA_2019_Str2) / 1000 
```

# Mesh sensitivity analysis 
Because the area have several islands and a complex coastline we will implement a barrier mesh. 
 
To optimize mesh characteristics, we followed the empirical recommendations of Zuur et al. (2017), Zuur & Ieno (2018) and H. Baka repository (https://haakonbakkagit.github.io/, accessed 30-09-2025). Specifically, 
 The cutoff distance should be approximately 1/5 of the max.edge of the inner area.
The max.edge can be approximated as 1/5 of an expected/guessed autocorrelation range.
The resolution of the outer area can be lower, with a general recommended inner-to-outer ratio of 1:5. 
Offset is recommended to extend to 2/3 of the expected range.

We run a sensitivity analysis to select the optimal mesh.

First we assume an expected/guessed autocorrelation range of 23nmi (42.5Km) and flowing the above recommendations we fit the M3 and 4 spatial models formulation

Except for theguessed autocorrelation range we fitted the same models with several meshes of higher and lower resolutions.
Models are compared using the AIC criterion and the mesh that resulted to the model fit with the lower AIC was selected as the one with the optimal mesh characteristics.

##BARIER MESHES for the Spatial Models
```{r}
###########################
#  Create the barrier mesh
###########################
create_mesh_Str1<- function(data,PA_polygone,edge_est_nmi)
{
df_name <- deparse(substitute(data))
mesh_name <- paste0("bspde", df_name,edge_est_nmi)

data <- data %>% filter(!is.na(Abundance))
data <- data %>% filter(!is.na(w_sn))
data.eggs <- data

poly.coast <- coast # Greek shp WGS84 
stations_sf <- st_as_sf(data.eggs, coords = c("Lon","Lat"), crs = 4326) 
plot(stations_sf$geometry, pch=19, col=2)

#Create a convex hull polygon around all stations 
transect_poly <- st_convex_hull(st_union(stations_sf))

# Buffer around the transect polygon (e.g., 0.1 degrees ~ ~10km depending on latitude)
buff <- st_buffer(transect_poly, dist = 10000) # WGS84 degrees; or use UTM for meters
plot(buff, col = "lightblue", border = "blue");plot(stations_sf$geometry, add = TRUE, pch = 19, col = 2)

output = st_intersection(buff,poly.coast)
plot(output,col=3)
domain = st_difference(transect_poly,st_union(output)) 
domain = st_sf(domain)

domain_utm <- st_transform(domain, "+init=epsg:32635 +units=km")
plot(domain_utm,col=4,axes=T);plot(data.eggs$X, data.eggs$Y, pch="+",add=T)

# Mesh characteristics, expected variogram range (~25 nm)
max.edge =edge_est_nmi 
range_est =max.edge*5 
bound.outer = range_est*(2/3)

#force geometries to be valid:
poly.coast <- st_make_valid(poly.coast)
domain_utm <- st_make_valid(domain_utm)

# Build high-resolution mesh
mesh_high = inla.mesh.2d( 
  boundary = domain_utm, 
  loc = cbind(data.eggs$X, data.eggs$Y),
  max.edge = c(1,5)*max.edge,
  cutoff = max.edge/5, 
  offset = c(max.edge, bound.outer)
  )

plot(mesh_high, lwd=0.5, axes = T); points(data.eggs$X, data.eggs$Y, col="red")
mesh_high <- make_mesh(data.eggs, c("X", "Y"), mesh = mesh_high) 
plot(mesh_high, lwd=0.5, axes = T); points(data.eggs$X, data.eggs$Y, col="red")

# Add on the barrier mesh component
poly.coast_utm <- st_transform(poly.coast, "+init=epsg:32635 +units=km")
bspde <- add_barrier_mesh(
  spde_obj = mesh_high,
  barrier_sf = poly.coast_utm,
  range_fraction = 0.1,
  proj_scaling = 1,
  plot = TRUE)
plot(bspde, lwd=0.5); points(data.eggs$X, data.eggs$Y, col="red")
assign(mesh_name, bspde, envir = .GlobalEnv)

}


create_mesh_Str2<- function(data,PA_polygone,edge_est_nmi)
{

df_name <- deparse(substitute(data))
mesh_name <- paste0("bspde", df_name,edge_est_nmi)

data <- data %>% filter(!is.na(Abundance))
data <- data %>% filter(!is.na(w_sn))
data.eggs <- data

poly.coast <- coast # Greek shp WGS84 
stations_sf <- st_as_sf(data.eggs, coords = c("Lon","Lat"), crs = 4326) 
plot(stations_sf$geometry, pch=19, col=2)

#Create a convex hull polygon around all stations 
transect_poly <- st_convex_hull(st_union(stations_sf))

# Buffer around the transect polygon (e.g., 0.1 degrees ~ ~10km depending on latitude)
buff <- st_buffer(transect_poly, dist = 10000) # WGS84 degrees; or use UTM for meters
plot(buff, col = "lightblue", border = "blue");plot(stations_sf$geometry, add = TRUE, pch = 19, col = 2)

output = st_intersection(buff,poly.coast)
plot(output,col=3)
domain = st_difference(transect_poly,st_union(output)) 
domain = st_sf(domain)

domain_utm <- st_transform(domain, "+init=epsg:32635 +units=km")
plot(domain_utm,col=4,axes=T);plot(data.eggs$X, data.eggs$Y, pch="+",add=T)

# Mesh characteristics, e
range_est<-  edge_est_nmi*5  
max.edge <- edge_est_nmi
bound.outer = range_est*(2/3)

#force geometries to be valid:
poly.coast <- st_make_valid(poly.coast)
domain_utm <- st_make_valid(domain_utm)

# Build high-resolution mesh
mesh_high = inla.mesh.2d( 
  boundary = domain_utm, 
  loc = cbind(data.eggs$X, data.eggs$Y),
  max.edge = c(1,5)*max.edge,
  cutoff = max.edge/5, 
  offset = c(max.edge, bound.outer)
  )

plot(mesh_high, lwd=0.5, axes = T); points(data.eggs$X, data.eggs$Y, col="red")
mesh_high <- make_mesh(data.eggs, c("X", "Y"), mesh = mesh_high) 
plot(mesh_high, lwd=0.5, axes = T); points(data.eggs$X, data.eggs$Y, col="red")

# Add on the barrier mesh component
poly.coast_utm <- st_transform(poly.coast, "+init=epsg:32635 +units=km")
bspde <- add_barrier_mesh(
  spde_obj = mesh_high,
  barrier_sf = poly.coast_utm,
  range_fraction = 0.1,
  proj_scaling = 1,
  plot = TRUE)

plot(bspde, lwd=0.5); points(data.eggs$X, data.eggs$Y, col="red")
assign(mesh_name, bspde, envir = .GlobalEnv)
}

```

## Code to fit spatial NB and Tweede model 
```{r}
fit_glm_NBsp_M3 <- function(data,PA_polygone,bspde) {
 
# Get data frame name
df_name <- deparse(substitute(data))
bspde_name <- deparse(substitute(bspde))
model_name <- paste0("model_M3", df_name,bspde_name)
Survey <- as.character(data$Survey[1])

data <- data %>% filter(!is.na(Abundance))
data <- data %>% filter(!is.na(w_sn))
Year <- as.character(data$Year[1])

# Fit model
model3 <-  sdmTMB(
  Ni ~ 1 + (Age),
  data = data,
  offset= log(data$Ri), # log because we are at the logit 
  family = nbinom2(link = "log"),
  mesh = bspde,
  spatial = "on")  

assign(model_name, model3, envir = .GlobalEnv)

#Summary
summary(model3)
sanity(model3)
}

fit_glm_TWsp_M4 <- function(data,PA_polygone,bspde) {
 
# Get data frame name
df_name <- deparse(substitute(data))
bspde_name <- deparse(substitute(bspde))
model_name <- paste0("model_M4", df_name,bspde_name)
Survey <- as.character(data$Survey[1])

data <- data %>% filter(!is.na(Abundance))
data <- data %>% filter(!is.na(w_sn))
Year <- as.character(data$Year[1])

# Fit model
model4 <-  sdmTMB(
  Abundance ~ 1 + (Age),
  data = data,
  family = tweedie(link = "log"),
  mesh = bspde,
  spatial = "on")  

assign(model_name, model4, envir = .GlobalEnv)

#Summary
summary(model4)
sanity(model4)
}
```
## Str1_2006 --> select edge_est_nmi=22  based on M3 & M4
### Fit the meshes
```{r}
create_mesh_Str1(Str1_06, PA_2006_Str1,edge_est_nmi=5)  
create_mesh_Str1(Str1_06, PA_2006_Str1,edge_est_nmi=7) 
create_mesh_Str1(Str1_06, PA_2006_Str1,edge_est_nmi=8.5) 
create_mesh_Str1(Str1_06, PA_2006_Str1,edge_est_nmi=10) 
create_mesh_Str1(Str1_06, PA_2006_Str1,edge_est_nmi=12) 
create_mesh_Str1(Str1_06, PA_2006_Str1,edge_est_nmi=15) 
create_mesh_Str1(Str1_06, PA_2006_Str1,edge_est_nmi=20)
create_mesh_Str1(Str1_06, PA_2006_Str1,edge_est_nmi=22)
create_mesh_Str1(Str1_06, PA_2006_Str1,edge_est_nmi=25)
create_mesh_Str1(Str1_06, PA_2006_Str1,edge_est_nmi=27)
```

### Fit the models
```{r}
fit_glm_NBsp_M3(Str1_06,PA_2006_Str1,bspdeStr1_065)   #Matern range too small
fit_glm_NBsp_M3(Str1_06,PA_2006_Str1,bspdeStr1_067) 
fit_glm_NBsp_M3(Str1_06,PA_2006_Str1,bspdeStr1_068.5) #Matern range too small
fit_glm_NBsp_M3(Str1_06,PA_2006_Str1,bspdeStr1_0610)
fit_glm_NBsp_M3(Str1_06,PA_2006_Str1,bspdeStr1_0612) 
fit_glm_NBsp_M3(Str1_06,PA_2006_Str1,bspdeStr1_0615)
fit_glm_NBsp_M3(Str1_06,PA_2006_Str1,bspdeStr1_0620)
fit_glm_NBsp_M3(Str1_06,PA_2006_Str1,bspdeStr1_0622)
fit_glm_NBsp_M3(Str1_06,PA_2006_Str1,bspdeStr1_0625)
fit_glm_NBsp_M3(Str1_06,PA_2006_Str1,bspdeStr1_0627)

fit_glm_TWsp_M4(Str1_06,PA_2006_Str1,bspdeStr1_065)  #Matern range too small
fit_glm_TWsp_M4(Str1_06,PA_2006_Str1,bspdeStr1_067)  
fit_glm_TWsp_M4(Str1_06,PA_2006_Str1,bspdeStr1_068.5) 
fit_glm_TWsp_M4(Str1_06,PA_2006_Str1,bspdeStr1_0610)
fit_glm_TWsp_M4(Str1_06,PA_2006_Str1,bspdeStr1_0612)
fit_glm_TWsp_M4(Str1_06,PA_2006_Str1,bspdeStr1_0615)
fit_glm_TWsp_M4(Str1_06,PA_2006_Str1,bspdeStr1_0620)
fit_glm_TWsp_M4(Str1_06,PA_2006_Str1,bspdeStr1_0622)
fit_glm_TWsp_M4(Str1_06,PA_2006_Str1,bspdeStr1_0625)
fit_glm_TWsp_M4(Str1_06,PA_2006_Str1,bspdeStr1_0627)
```

### Compare models and decide the edge
```{r}
modelsM3 <- list(m3_5=      model_M3Str1_06bspdeStr1_065,
                 m3_7=      model_M3Str1_06bspdeStr1_067,   
                 m3_8.5 =   model_M3Str1_06bspdeStr1_068.5, 
                 m3_10=     model_M3Str1_06bspdeStr1_0610,  
                 m3_12=     model_M3Str1_06bspdeStr1_0612, 
                 m3_15=     model_M3Str1_06bspdeStr1_0615,
                 m3_20=     model_M3Str1_06bspdeStr1_0620,
                 m3_22=     model_M3Str1_06bspdeStr1_0622,
                 m3_25=     model_M3Str1_06bspdeStr1_0625,
                 m3_27=     model_M3Str1_06bspdeStr1_0627)

modelsM4 <- list(m4_5=      model_M4Str1_06bspdeStr1_065,
                 m4_7=      model_M4Str1_06bspdeStr1_067,   
                 m4_8.5 =   model_M4Str1_06bspdeStr1_068.5, 
                 m4_10=     model_M4Str1_06bspdeStr1_0610,  
                 m4_12=     model_M4Str1_06bspdeStr1_0612, 
                 m4_15=     model_M4Str1_06bspdeStr1_0615,
                 m4_20=     model_M4Str1_06bspdeStr1_0620,
                 m4_22=     model_M4Str1_06bspdeStr1_0622,
                 m4_25=     model_M4Str1_06bspdeStr1_0625,
                 m4_27=     model_M4Str1_06bspdeStr1_0627)

AIC_dfM3 <- tibble(
  model = names(modelsM3),
  AIC = sapply(modelsM3, AIC)
) %>% mutate(ΔAIC = AIC - min(AIC))
AIC_dfM3<- AIC_dfM3 %>% filter(model %in% c("m3_5","m3_7","m3_8.5", "m3_10","m3_12", "m3_15", "m3_20","m3_22","m3_25","m3_27")) %>% mutate(ΔAIC = AIC - min(AIC))
AIC_dfM3

AIC_dfM4 <- tibble(
  model = names(modelsM4),
  AIC = sapply(modelsM4, AIC)
) %>% mutate(ΔAIC = AIC - min(AIC))
AIC_dfM4<- AIC_dfM4 %>% filter(model %in% c("m4_5","m4_7","m4_8.5", "m4_10","m4_12", "m4_15", "m4_20","m4_22","m4_25","m4_27")) %>% mutate(ΔAIC = AIC - min(AIC))
AIC_dfM4
```


## Str2_2006 --> select edge_est_nmi= 7   based on M3 & M4
### Fit the meshes
```{r}
create_mesh_Str2(Str2_06, PA_2006_Str2,edge_est_nmi=5) 
create_mesh_Str2(Str2_06, PA_2006_Str2,edge_est_nmi=7) 
create_mesh_Str2(Str2_06, PA_2006_Str2,edge_est_nmi=8.5) 
create_mesh_Str2(Str2_06, PA_2006_Str2,edge_est_nmi=10) 
create_mesh_Str2(Str2_06, PA_2006_Str2,edge_est_nmi=12) 
create_mesh_Str2(Str2_06, PA_2006_Str2,edge_est_nmi=15) 
create_mesh_Str2(Str2_06, PA_2006_Str2,edge_est_nmi=20)
create_mesh_Str2(Str2_06, PA_2006_Str2,edge_est_nmi=22)
```

### Fit the models
```{r}
fit_glm_NBsp_M3(Str2_06,PA_2006_Str2,bspdeStr2_065)   
fit_glm_NBsp_M3(Str2_06,PA_2006_Str2,bspdeStr2_068.5)
fit_glm_NBsp_M3(Str2_06,PA_2006_Str2,bspdeStr2_0610)
fit_glm_NBsp_M3(Str2_06,PA_2006_Str2,bspdeStr2_0612)  
fit_glm_NBsp_M3(Str2_06,PA_2006_Str2,bspdeStr2_0615) 
fit_glm_NBsp_M3(Str2_06,PA_2006_Str2,bspdeStr2_0620)
fit_glm_NBsp_M3(Str2_06,PA_2006_Str2,bspdeStr2_0622)

fit_glm_TWsp_M4(Str2_06,PA_2006_Str2,bspdeStr2_065)  
fit_glm_TWsp_M4(Str2_06,PA_2006_Str2,bspdeStr2_067)  
fit_glm_TWsp_M4(Str2_06,PA_2006_Str2,bspdeStr2_068.5) 
fit_glm_TWsp_M4(Str2_06,PA_2006_Str2,bspdeStr2_0610)
fit_glm_TWsp_M4(Str2_06,PA_2006_Str2,bspdeStr2_0612)  
fit_glm_TWsp_M4(Str2_06,PA_2006_Str2,bspdeStr2_0615)  
fit_glm_TWsp_M4(Str2_06,PA_2006_Str2,bspdeStr2_0620)
fit_glm_TWsp_M4(Str2_06,PA_2006_Str2,bspdeStr2_0622)
```

### Compare models and decide the edge
```{r}
modelsM3 <- list(m3_5=      model_M3Str2_06bspdeStr2_065,
                 m3_7=      model_M3Str2_06bspdeStr2_067,   
                 m3_8.5 =   model_M3Str2_06bspdeStr2_068.5, 
                 m3_10=     model_M3Str2_06bspdeStr2_0610,  
                 m3_12=     model_M3Str2_06bspdeStr2_0612, 
                 m3_15=     model_M3Str2_06bspdeStr2_0615,
                 m3_20=     model_M3Str2_06bspdeStr2_0620,
                 m3_22=     model_M3Str2_06bspdeStr2_0622)

modelsM4 <- list(m4_5=      model_M4Str2_06bspdeStr2_065,
                 m4_7=      model_M4Str2_06bspdeStr2_067,   
                 m4_8.5 =   model_M4Str2_06bspdeStr2_068.5, 
                 m4_10=     model_M4Str2_06bspdeStr2_0610,  
                 m4_12=     model_M4Str2_06bspdeStr2_0612, 
                 m4_15=     model_M4Str2_06bspdeStr2_0615,
                 m4_20=     model_M4Str2_06bspdeStr2_0620,
                 m4_22=     model_M4Str2_06bspdeStr2_0622)

AIC_dfM3 <- tibble(
  model = names(modelsM3),
  AIC = sapply(modelsM3, AIC)
) %>% mutate(ΔAIC = AIC - min(AIC))
AIC_dfM3<- AIC_dfM3 %>% filter(model %in% c("m3_5","m3_7","m3_8.5", "m3_10","m3_12", "m3_15", "m3_20", "m3_22")) %>% mutate(ΔAIC = AIC - min(AIC))
AIC_dfM3

AIC_dfM4 <- tibble(
  model = names(modelsM4),
  AIC = sapply(modelsM4, AIC)
) %>% mutate(ΔAIC = AIC - min(AIC))
AIC_dfM4<- AIC_dfM4 %>% filter(model %in% c("m4_5","m4_7","m4_8.5", "m4_10","m4_12", "m4_15", "m4_20","m4_22")) %>% mutate(ΔAIC = AIC - min(AIC))
AIC_dfM4
```


