---
title: "Po,z,Ptot estimation"
author: "Eleni Servou"
date: todays
format:
  html:
    toc: true
    toc-depth: 4
    toc-float: true
    code-fold: true         
    code-summary: "Hide/Show "
    code-tools:
      toggle: true         
---

## Load libraries and prepare data
```{r}
library(readxl)
library(readr)
library(ggplot2)
library(dplyr)
library(glmmTMB)
library(sdmTMB)
library(cowplot)
library(gt)
library(DHARMa)
library(sf)
library(INLA)
library(sdmTMBextra)
library(writexl)
library(scales)
library(tidyr)
library(stringr)
library(viridis)
library(patchwork)
library(RColorBrewer)
library(gstat)
```

## Load egg data
```{r}
# Eastern stratum (2006)
Str1_06 <- read_xlsx("data.xlsx", sheet = "2006 east")
Str1_06$Survey <- "Str1_06"

# Western stratum (2006)
Str2_06 <- read_xlsx("data.xlsx", sheet = "2006 west")
Str2_06$Survey <- "Str2_06"
```


## For mapping
```{r}
# Load coastline data for Greece
coast <- st_read("Greece_geology/Greece/land_poly_GR.gpkg", quiet = TRUE) %>% 
  st_transform(crs = "WGS84")

# Transform coordinates to km (UTM Zone 35N)
coast_km <- st_transform(coast, crs = 32635)
st_geometry(coast_km) <- st_geometry(coast_km) / 1000## For mapping

# Load coastline data for Greece
coast <- st_read("Greece_geology/Greece/land_poly_GR.gpkg", quiet = TRUE) %>% 
  st_transform(crs = "WGS84")

# Transform coordinates to km (UTM Zone 35N)
coast_km <- st_transform(coast, crs = 32635)
st_geometry(coast_km) <- st_geometry(coast_km) / 1000
```

## Voronoi polygons of A1 
```{r}
# Read shapefiles and transform to WGS84
PA_2006_Str1 <- st_read("voronois/str1_sd_2006_dis.shp") %>% 
  st_transform(crs = "WGS84")

PA_2006_Str2 <- st_read("voronois/str2_sd_2006_dis.shp") %>% 
  st_transform(crs = "WGS84")

# Transform coordinates to kilometers (UTM Zone 35N) - Eastern stratum
PA_2006_Str1 <- PA_2006_Str1 %>% 
  st_transform(crs = 32635)
st_geometry(PA_2006_Str1) <- st_geometry(PA_2006_Str1) / 1000

# Transform coordinates to kilometers (UTM Zone 35N) - Western stratum
PA_2006_Str2 <- PA_2006_Str2 %>% 
  st_transform(crs = 32635)
st_geometry(PA_2006_Str2) <- st_geometry(PA_2006_Str2) / 1000


# Extract total spawning area (converted to m²)
area_Str1 <- c(as.numeric(PA_2006_Str1$Area)*10^6)
               
area_Str2 <- c(as.numeric(PA_2006_Str2$Area)*10^6)
```

## Estimate station weights
```{r}
# Calculate weights for Eastern stratum 
Str1_06_w <- Str1_06 %>%
  distinct(Station, .keep_all = TRUE) %>%
  mutate(total_km = sum(km_2, na.rm = TRUE)) %>%
  mutate(w_sn = km_2 / mean(km_2, na.rm = TRUE))

Str1_06 <- Str1_06 %>%
  left_join(Str1_06_w %>% select(Station, w_sn), by = "Station") %>%
  add_utm_columns(
    ll_names = c("Lon", "Lat"),
    ll_crs = "WGS84",
    utm_crs = 32635,
    utm_names = c("X", "Y"),
    units = "km")

# Calculate weights for Western stratum
Str2_06_w <- Str2_06 %>%
  distinct(Station, .keep_all = TRUE) %>%
  mutate(total_km = sum(km_2, na.rm = TRUE)) %>%
  mutate(w_sn = km_2 / mean(km_2, na.rm = TRUE))

Str2_06 <- Str2_06 %>%
  left_join(Str2_06_w %>% select(Station, w_sn), by = "Station") %>%
  add_utm_columns(
    ll_names = c("Lon", "Lat"),
    ll_crs = "WGS84",
    utm_crs = 32635,
    utm_names = c("X", "Y"),
    units = "km"
  )


```


# Model 0: Weighted non-linear least squares regression model

The classic weighted exponential model of Picquelle and Stauffer 1985 used in Somarakis et al. 2012

### Write M0 function 
```{r}
#Crate a function that will do everything at once

fit_M0_model_M0 <- function(data) {
# Get dataframe name
  df_name <- deparse(substitute(data))
  model_name <- paste0("model_M0", df_name)
  Survey <- as.character(data$Survey[1])
  
# Build dynamic object names
  plot_name <- paste0("plot_M0_", df_name)
  plot_log_name <- paste0("plot_log_M0_", df_name)
  df_pred_name <- paste0("pred_M0_", df_name)
  P_est_name <- paste0("P_est_M0_", df_name)
  z_est_name <- paste0("z_est_M0_", df_name)
  
  P_se_name <- paste0("P_se_M0_", df_name)
  z_se_name <- paste0("z_se_M0_", df_name)
  
  P_ci_name <- paste0("P_ci_M0_", df_name)
  z_ci_name <- paste0("z_ci_M0_", df_name)
  
# Set a range of initial values and fit model
  model0 <- nls.multstart::nls_multstart(
    Abundance ~ P * exp(-z * Age),
    data = data,
    iter = 250,
    convergence_count = FALSE,
    start_lower = c(P = 50, z = 0.1),
    start_upper = c(P = 200, z = 0.8),
    modelweights = data$w_sn,
    na.action = na.omit
  )
  
assign(model_name, model0, envir = .GlobalEnv)

# Summary
  m <- summary(model0)
  print(m)
  
  # Extract params
  params <- coef(model0)
  P_est <- params["P"]
  z_est <- params["z"]
  
  # SEs
  P_se <- m$parameters["P", "Std. Error"]
  z_se <- m$parameters["z", "Std. Error"]
  
  # CIs
  ci <- confint(model0)
  P_ci <- ci["P", ]
  z_ci <- ci["z", ]
  
# Output tables
  outablePo <- data.frame(
    dataset   = c(Survey),
    parameter = c("P"),
    estimate  = c(P_est),
    se        = c(P_se),
    ci_lower  = c(P_ci[1]),
    ci_upper  = c(P_ci[2]),
    CV        = c(P_se / P_est),
    model     = "M0",
    stringsAsFactors = FALSE
  )
  
  outableZ <- data.frame(
    dataset   = c(Survey),
    parameter = c("z"),
    estimate  = c(z_est),
    se        = c(z_se),
    ci_lower  = c(z_ci[1]),
    ci_upper  = c(z_ci[2]),
    CV        = c(z_se / z_est),
    model     = "M0",
    stringsAsFactors = FALSE
  )
  
# Generate predictions for plotting
  x_new <- seq(
    min(data$Age, na.rm = TRUE),
    max(data$Age, na.rm = TRUE),
    length.out = 10000
  )
  
  y_pred <- predict(model0, newdata = data.frame(Age = x_new))
  df_pred <- data.frame(x = x_new, y = y_pred)

# Plot
  plot <- ggplot(data, aes(x = Age, y = Abundance)) +
    geom_point(color = "black", alpha = 0.6) +
    geom_line(data = df_pred, aes(x, y), color = brewer.pal(7, "Dark2")[1], size = 1) +
    labs(
      title = "",
      subtitle = paste(df_name),
      x = "Age",
      y = "Abundance"
    ) +
    theme_bw() +
    theme(panel.grid.major = element_line(color = "white"))
  
  assign(plot_name, plot, envir = .GlobalEnv)
  
  return(list(outablePo = outablePo, outableZ = outableZ))
}
```

### Fit M0
```{r}
# Use the function to fit all models together
results_all_M0 <- bind_rows(
  fit_M0_model_M0(Str1_06),
  fit_M0_model_M0(Str2_06)
)
```

### Get P, z results and estimate Ptot
```{r}
# Filter and format Po results for Eastern stratum (Str1)
Str1_resultsPo <- results_all_M0$outablePo %>% 
  filter(grepl("^Str1_", dataset)) %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3)))
gt(Str1_resultsPo)

# Filter and format Po results for Western stratum (Str2)
Str2_resultsPo <- results_all_M0$outablePo %>% 
  filter(grepl("^Str2_", dataset)) %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3)))
gt(Str2_resultsPo)

# Filter and format Z results for Eastern stratum (Str1)
Str1_resultsZ <- results_all_M0$outableZ %>% 
  filter(grepl("^Str1_", dataset)) %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3)))
gt(Str1_resultsZ)

# Filter and format Z results for Western stratum (Str2)
Str2_resultsZ <- results_all_M0$outableZ %>% 
  filter(grepl("^Str2_", dataset)) %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3)))
gt(Str2_resultsZ)

# Estimate Ptot for Eastern stratum (Str1)
Str1_resultsPtot <- tibble(
  dataset   = Str1_resultsPo$dataset,
  parameter = "Ptot",
  estimate  = Str1_resultsPo$estimate * area_Str1,
  se        = NA,
  ci_lower  = Str1_resultsPo$ci_lower * area_Str1,
  ci_upper  = Str1_resultsPo$ci_upper * area_Str1,
  CV        = Str1_resultsPo$CV,
  model     = Str1_resultsPo$model
) %>%
  mutate(se = estimate * CV)

# Estimate Ptot for Western stratum (Str2)
Str2_resultsPtot <- tibble(
  dataset   = Str2_resultsPo$dataset,
  parameter = "Ptot",
  estimate  = Str2_resultsPo$estimate * area_Str2,
  se        = NA,
  ci_lower  = Str2_resultsPo$ci_lower * area_Str2,
  ci_upper  = Str2_resultsPo$ci_upper * area_Str2,
  CV        = Str2_resultsPo$CV,
  model     = Str2_resultsPo$model
) %>%
  mutate(se = estimate * CV)
```

### Summary plots
```{r}
# Eastern stratum Po plot
plot_Po_Str1 <- ggplot() +
  geom_point(data = Str1_resultsPo,
             aes(y = estimate, x = dataset, col = model),
             size = 2) +
  geom_errorbar(data = Str1_resultsPo, 
                aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model),
                width = 0.2, linewidth = 1) +
   theme_bw() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
   ylab(expression(P[0] ~ (eggs ~ m^-2 ~ day^-1))) +
   xlab("") +
   scale_fill_manual(name = "Model", values = brewer.pal(1, "Set1"))

# Western stratum Po plot
plot_Po_Str2 <- ggplot() +
  geom_point(data = Str2_resultsPo,
             aes(y = estimate, x = dataset, col = model),
             size = 2) +
  geom_errorbar(data = Str2_resultsPo, 
                aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model), 
                width = 0.2, linewidth = 1) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[0] ~ (eggs ~ m^-2 ~ day^-1))) +
  xlab("") +
  scale_fill_manual(name = "Model", values = brewer.pal(1, "Set1"))

# Eastern stratum Z plot
plot_z_Str1 <- ggplot() +
  geom_point(data = Str1_resultsZ,
             aes(y = estimate, x = dataset, col = model), 
             size = 2) +
  geom_errorbar(data = Str1_resultsZ,
                aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model),
                width = 0.2, linewidth = 1) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(Z ~ (day^-1))) +
  xlab("") +
  scale_fill_manual(name = "Model", values = brewer.pal(1, "Set1"))

# Western stratum Z plot
plot_z_Str2 <- ggplot() +
  geom_point(data = Str2_resultsZ, 
             aes(y = estimate, x = dataset, col = model),
             size = 2) +
  geom_errorbar(data = Str2_resultsZ, 
                aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model), 
                width = 0.2, linewidth = 1) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(Z ~ (day^-1))) +
  xlab("") +
  scale_fill_manual(name = "Model", values = brewer.pal(1, "Set1"))

# Eastern stratum Ptot plot
plot_Ptot_Str1 <- ggplot() +
  geom_point(data = Str1_resultsPtot, 
             aes(y = estimate / 10^11, x = dataset, col = model),
             size = 2) +
  geom_errorbar(data = Str1_resultsPtot,
                aes(x = dataset, ymin = ci_lower / 10^11, ymax = ci_upper / 10^11, color = model),
                width = 0.2, linewidth = 1) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[tot] ~ (eggs ~ day^-1) ~ x10^11)) +
  xlab("") +
  scale_fill_manual(name = "Model", values = brewer.pal(1, "Set1"))

# Western stratum Ptot plot
plot_Ptot_Str2 <- ggplot() +
  geom_point(data = Str2_resultsPtot, 
             aes(y = estimate / 10^11, x = dataset, col = model), 
             size = 2) +
  geom_errorbar(data = Str2_resultsPtot, 
                aes(x = dataset, ymin = ci_lower / 10^11, ymax = ci_upper / 10^11, color = model), 
                width = 0.2, linewidth = 1) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[tot] ~ (eggs ~ day^-1) ~ x10^11)) +
  xlab("") +
  scale_fill_manual(name = "Model", values = brewer.pal(1, "Set1"))



# Combine plots into a grid layout
Summary_Plots <- plot_grid(
  plot_Po_Str1, plot_Po_Str2,
  plot_z_Str1, plot_z_Str2,
  plot_Ptot_Str1, plot_Ptot_Str2,
  ncol = 2)
print(Summary_Plots)
```

### Visualize Fits 
```{r}
#Plots of fitted models
Ploted_fits_Str1<- print(plot_grid(plot_M0_Str1_06))
Ploted_fits_Str2<- print(plot_grid(plot_M0_Str2_06))
```

### QQ and Residual
```{r}
Residuals_plots <- function(data, model, model_name) {
  df_name <- deparse(substitute(data))
  res_plot_name <- paste0("res_plot_", model_name, df_name)
  res_qq_name <- paste0("res_qq_", model_name, df_name)
  
  data <- data %>% filter(!is.na(Abundance), !is.na(w_sn))
  year <- sub(".*_(\\d{2}).*", "\\1", df_name)
  year <- paste0("20", year)
  stratum <- case_when(
    grepl("^Str1", df_name) ~ "Eastern stratum",
    grepl("^Str2", df_name) ~ "Western stratum",
    TRUE ~ ""
  )
#Residual  
 df_res <- data.frame(
    Age = data$Age,
    Predicted = predict(model, newdata = data.frame(Age = data$Age)),
    Observed = data$Abundance
  ) %>% mutate(Residual = (residuals(model)) / summary(model)$sigma)
  
# Pearson Cor
  r_val <- cor(df_res$Predicted, df_res$Observed, use =  "complete.obs",method = "pearson")
  
# Residual vs Fitted plot
p_obs_pred <- ggplot(df_res, aes(x = Predicted, y = Observed)) +
    geom_point(alpha = 0.6, color = brewer.pal(7, "Dark2")[1]) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
    labs(
      x = "",
      y = "Observed",
      subtitle = paste(model_name)
    ) +
    annotate(
      "text",
      x = -Inf,
      y = Inf,
      label = sprintf("r = %.3f", r_val),
      hjust = -0.1,
      vjust = 1.1,
      size = 4, color = "red"
    ) +
    theme_bw() +
    theme(plot.title = element_text(size = 11),plot.margin = margin(t = 0.2, r = 1, b = 0, l = 1))
  
  p_res_fit <- ggplot(df_res, aes(x = Predicted, y = Residual)) +
    geom_point(alpha = 0.6, color = brewer.pal(7, "Dark2")[1]) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    labs(
      x = "",
      y = "Residuals",
      subtitle = ""
    ) +
    theme_bw() +
    theme(plot.margin = margin(t = 0.2, r = 1, b = 0, l = 1))
  
  p_qq <- ggplot(df_res, aes(sample = Residual)) +
    stat_qq(color = brewer.pal(7, "Dark2")[1]) +
    stat_qq_line(linetype = "dashed", color = "red") +
    labs(
      x = "",
      y = "Sample Q",
      subtitle = ""
    ) + 
    theme_bw() +
    theme(plot.margin = margin(t = 0.2, r = 1, b = 0, l = 1))
  
  res_plots <- plot_grid(p_obs_pred, p_res_fit, p_qq, ncol = 3)
  assign(res_plot_name, res_plots, envir = .GlobalEnv)
  assign(res_qq_name, p_qq, envir = .GlobalEnv)
}

Residuals_plots(Str1_06, model_M0Str1_06, "M0")
res_plot_M0Str1_06
res_qq_M0Str1_06

Residuals_plots(Str2_06, model_M0Str2_06, "M0")
res_plot_M0Str2_06
res_qq_M0Str2_06
```

# Model 1: GLM with NB2 with OFFSET(Ri) - without spatial autocorrelation

Let's move to GLM NB2.
I have tested the results from `glmmTMB`, `sdmTMB`, and `glm.nb` from `MASS` — all give the exact same output.

### Write M1 function
```{r}
fit_glm_NB_M1 <- function(data,plot_basic) {

# Get data frame name
df_name <- deparse(substitute(data))
model_name <- paste0("model_M1", df_name)
Survey <- as.character(data$Survey[1])

data <- data %>% filter(!is.na(Abundance))
data <- data %>% filter(!is.na(w_sn))
Year <- as.character(data$Year[1])

# Build dynamic object names
  plot_name  <- paste0("plot_M1_", df_name)
  plot_log_name  <- paste0("plot_log_M1_", df_name)
  dispersion_name  <- paste0("dispersion_M1_", df_name)

  P_est_name <- paste0("P_est_M1_", df_name)
  z_est_name <- paste0("z_est_M1_", df_name)
  
  P_se_name  <- paste0("P_se_M1_", df_name)
  z_se_name  <- paste0("z_se_M1_", df_name)
  
  P_ci_name  <- paste0("P_ci_M1_", df_name)
  z_ci_name  <- paste0("z_ci_M1_", df_name)
  
# Fit model
model1 <- sdmTMB(
  Ni ~ 1 + (Age),
  data = data,
  offset= log(data$Ri), 
  family = nbinom2(link = "log"),
  weights = data$w_sn,
  spatial = "off")  

assign(model_name, model1, envir = .GlobalEnv)
  
#Summary/diagnostics
summary(model1)
sanity(model1)
  
params <-tidy(model1, conf.int = TRUE)
tidy(model1, effects = "ran_par", conf.int = TRUE)

# Extract param
 P_est <- exp(params$estimate[1]) 
 z_est <- -(params$estimate[2]) 
 
# SE
  P_se_log <-  params$std.error[1]; 
  z_se <- params$std.error[2];
  
# CI 
  P_ci <- NA;  
  z_ci <-  c(-params$conf.high[2], -params$conf.low[2]) 
  
# Compute transformed estimates
#LogP0 
CV_P <- sqrt(exp(P_se_log^2) - 1)
P_se <- P_est * CV_P
P_ci_lower <- exp(params$conf.low[1])  
P_ci_upper <- exp(params$conf.high[1])

outablePo <- data.frame(
   dataset       =  Survey,
   parameter     = c("P"),
   estimate_log  = params$estimate[1], 
   estimate      = c(P_est),       # Unlog
   se            = c(P_se) ,       #se of the estimated model,
   ci_lower      = c(P_ci_lower),
   ci_upper      = c(P_ci_upper),
   CV            = c(CV_P),       
   model         = "M1",
   stringsAsFactors = FALSE) 

outableZ <- data.frame(
    dataset   = c(Survey),
    parameter = "z",
    estimate  = c(z_est),
    se        = c(z_se),
    ci_lower  = c(z_ci[1]), #It haves been reversed because z insted of -z is modeld
    ci_upper  = c(z_ci[2]),
    CV        = c(z_se/z_est),
    model     = "M1",
    stringsAsFactors = FALSE)


dispersion <-model1$sd_report$value["phi"]%>% as.numeric()
assign(dispersion_name, dispersion, envir = .GlobalEnv)

#Generate predictions for plotting
 # To predict we need an offsett so we will just predict in the original data
  y_pred <- predict(model1, newdata = data.frame(Age = data$Age),se_fit = TRUE, re_form = NA) 
  
  #offset = log(data$Ri)) # unlog the prediction
  df_pred <- data.frame(x = data$Age, y = exp(y_pred$est),
                        ymin = exp((y_pred$est - 1.96 * (y_pred$est_se))),
                        ymax = exp((y_pred$est + 1.96 * (y_pred$est_se)))) #If i want to plot te CI in the model 
# Plot
  plot <- 
     plot_basic +
   geom_line(data = df_pred, aes(x, y), color = brewer.pal(7, "Dark2")[2], size = 1) +
   labs(
      title = "",
      subtitle = paste(df_name),
      x = "Age",
      y = "Abundance") +
    theme_bw() +
        theme(panel.grid.major = element_line(color = "white"))
  
assign(plot_name, plot, envir = .GlobalEnv) 
return(list(outablePo= outablePo, outableZ=outableZ))
}
```

### Fit M1
```{r}
results_all_M1 <- bind_rows(
  fit_glm_NB_M1(Str1_06, plot_M0_Str1_06),     
  fit_glm_NB_M1(Str2_06, plot_M0_Str2_06))
```

### Get P,z results and estimate Ptot
```{r}
outablePo_temp_Str1<- results_all_M1$outablePo %>% filter(grepl("^Str1_",dataset)) %>% mutate(across(where(is.numeric), ~ round(.x, 3)))
outablePo_temp_Str1 <- outablePo_temp_Str1 %>% select (-estimate_log)
Str1_resultsPo <- rbind(Str1_resultsPo,outablePo_temp_Str1)
gt(Str1_resultsPo)

outablePo_temp_Str2<- results_all_M1$outablePo %>% filter(grepl("^Str2",dataset)) %>% mutate(across(where(is.numeric), ~ round(.x, 3)))
outablePo_temp_Str2 <- outablePo_temp_Str2 %>% select (-estimate_log)
Str2_resultsPo <- rbind(Str2_resultsPo,outablePo_temp_Str2)
gt(Str2_resultsPo)

outableZ_temp_Str1 <- results_all_M1$outableZ %>% filter(grepl("^Str1_",dataset)) %>% mutate(across(where(is.numeric), ~ round(.x, 3)))
Str1_resultsZ <- rbind(Str1_resultsZ,outableZ_temp_Str1)
gt(Str1_resultsZ)

outableZ_temp_Str2 <- results_all_M1$outableZ %>% filter(grepl("^Str2_",dataset)) %>% mutate(across(where(is.numeric), ~ round(.x, 3)))
Str2_resultsZ <- rbind(Str2_resultsZ,outableZ_temp_Str2)
gt(Str2_resultsZ)


Str1_resultsPtot_temp <- tibble(
  dataset   = outablePo_temp_Str1$dataset,
  parameter = "Ptot",
  estimate  = outablePo_temp_Str1$estimate * area_Str1,
  se        = NA,
  ci_lower  = outablePo_temp_Str1$ci_lower * area_Str1,
  ci_upper  = outablePo_temp_Str1$ci_upper * area_Str1,
  CV        = outablePo_temp_Str1$CV,
  model     = outablePo_temp_Str1$model
) %>%
  mutate(se = estimate * CV)

Str1_resultsPtot <- rbind(Str1_resultsPtot,Str1_resultsPtot_temp)

Str2_resultsPtot_temp <- tibble(
  dataset   = outablePo_temp_Str2$dataset,
  parameter = "Ptot",
  estimate  = outablePo_temp_Str2$estimate * area_Str2,
  se        = NA,
  ci_lower  = outablePo_temp_Str2$ci_lower * area_Str2,
  ci_upper  = outablePo_temp_Str2$ci_upper * area_Str2,
  CV        = outablePo_temp_Str2$CV,
  model     = outablePo_temp_Str2$model
) %>%
  mutate(se = estimate * CV)

Str2_resultsPtot <- rbind(Str2_resultsPtot,Str2_resultsPtot_temp)
```

### Summary plots
```{r}
# Eastern stratum Po plot
plot_Po_Str1 <- ggplot() +
  geom_point(data = Str1_resultsPo, 
             aes(y = estimate, x = dataset, col = model),
             size = 2, position = position_dodge(width = 0.3)) + 
  geom_errorbar(data = Str1_resultsPo, 
                aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model),
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +  
  theme_bw() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[0] ~ (eggs ~ m^-2 ~ day^-1))) + xlab("") +
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))


# Western stratum Po plot
plot_Po_Str2 <-ggplot() +
  geom_point(data = Str2_resultsPo, 
             aes(y = estimate, x = dataset, col = model),
             size = 2, position = position_dodge(width = 0.3)) + 
  geom_errorbar(data = Str2_resultsPo, 
                aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model), 
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +  
  theme_bw() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[0] ~ (eggs ~ m^-2 ~ day^-1))) + xlab("") +
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))

# Eastern stratum Z plot
plot_z_Str1 <-
ggplot() +
  geom_point(data = Str1_resultsZ, 
             aes(y = estimate, x = dataset, col = model),
             size = 2, position = position_dodge(width = 0.3)) + 
  geom_errorbar(data = Str1_resultsZ, 
                aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model), 
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +  
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(Z ~ (day^-1))) + xlab("") +
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))

# Western stratum Z plot
plot_z_Str2 <-
ggplot() +
  geom_point(data = Str2_resultsZ,
             aes(y = estimate, x = dataset, col = model),
             size = 2, position = position_dodge(width = 0.3)) + 
  geom_errorbar(data = Str2_resultsZ, 
                aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model), 
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +  
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(Z ~ (day^-1))) + xlab("") +
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))

# Eastern stratum Ptot plot
plot_Ptot_Str1 <-ggplot() +
  geom_point(data = Str1_resultsPtot, 
             aes(y = estimate / 10^11, x = dataset, col = model), 
             size = 2, position =position_dodge(width = 0.3)) + 
  geom_errorbar(data = Str1_resultsPtot, 
                aes(x = dataset, ymin = ci_lower / 10^11, ymax = ci_upper / 10^11, color = model),
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +   
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[tot] ~ (eggs ~ day^-1) ~ x10^11)) + xlab("") +
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))

# Western stratum Ptot plot
plot_Ptot_Str2 <-ggplot() +
  geom_point(data = Str2_resultsPtot, 
             aes(y = estimate / 10^11, x = dataset, col = model),
             size = 2, position = position_dodge(width = 0.3)) + 
  geom_errorbar(data = Str2_resultsPtot, 
                aes(x = dataset, ymin = ci_lower / 10^11, ymax = ci_upper / 10^11, color = model),
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +   
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[tot] ~ (eggs ~ day^-1) ~ x10^11)) + xlab("") +
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))

Summary_Plots <-  
  print(plot_grid(
    plot_Po_Str1, plot_Po_Str2,
    plot_z_Str1, plot_z_Str2, 
    plot_Ptot_Str1, plot_Ptot_Str2,
    ncol = 2))
```

### Visualize Fits 
```{r}
#Plots of fitted models
Ploted_fits_Str1<- print(plot_grid(plot_M1_Str1_06))
Ploted_fits_Str2<- print(plot_grid(plot_M1_Str2_06))
```

### QQ and Residual
```{r}
Residuals_plots<- function(data,model,model_name){
    df_name <- deparse(substitute(data))
    res_plot_name  <- paste0("res_plot_",model_name,df_name)
    res_qq_name  <- paste0("res_qq_",model_name,df_name)
    data <- data %>% filter(!is.na(Abundance), !is.na(w_sn))
    year <- sub(".*_(\\d{2}).*", "\\1", df_name) ;year <- paste0("20", year)
   
 # Determine stratum
 stratum <- case_when(
  grepl("^Str1", df_name) ~ "Eastern stratum",
  grepl("^Str2", df_name) ~ "Western stratum",
  TRUE ~ "")

# Generate prediction
df_res <- data.frame(
  Age = data$Age,
  Predicted = exp((predict(model, newdata = data.frame(Age = data$Age)))$est), # Unlog
  Observed= data$Abundance) %>% mutate (Residual= residuals(model,type ="mle-mvn"))

#Pearson Cor 
r_val <- cor(df_res$Predicted, df_res$Observed, use = "complete.obs",method = "pearson")

# Residual vs Fitted plot
p_obs_pred <- ggplot(df_res, aes(x = Predicted, y = Observed)) +
  geom_point(alpha = 0.6,color=brewer.pal(7, "Dark2")[2]) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(
    x = "",
    y = "Observed",
    subtitle = paste(model_name) )+
  annotate(
  "text",
  x = -Inf,
  y = Inf,
  label = sprintf("r = %.3f", r_val),
  hjust = -0.1,
  vjust = 1.1,
  size =4, color="red") +
  theme_bw() +
  theme(plot.margin = margin(t=0,r = 1,b =0,l = 1))
   
p_res_fit <- ggplot(df_res, aes(x = Predicted, y = Residual)) +
    geom_point(alpha = 0.6,color=brewer.pal(7, "Dark2")[2]) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    theme_bw() +
    labs(
      x = "",
      y = "Residuals",
      subtitle="") +
    theme(plot.margin = margin(t=0,r = 1,b =0,l = 1))


p_qq <- ggplot(df_res, aes(sample = Residual)) +
    stat_qq(color=brewer.pal(7, "Dark2")[2]) +
    stat_qq_line(linetype = "dashed", color = "red") +
    theme_bw() +
    labs(
      x = "",
      y = "Sample Q",
      subtitle="") +
    theme(plot.margin = margin(t=0,r = 1,b =0,l = 1))


res_plots<- plot_grid(p_obs_pred,p_res_fit,p_qq, ncol=3)
assign(res_plot_name, res_plots, envir = .GlobalEnv)
assign(res_qq_name, p_qq, envir = .GlobalEnv)  
}

Residuals_plots(Str1_06,model_M1Str1_06, "M1") ; res_plot_M1Str1_06 ;res_qq_M1Str1_06
Residuals_plots(Str2_06,model_M1Str2_06, "M1") ; res_plot_M1Str2_06 ;res_qq_M1Str2_06

Residuals_Str1_06<- print(plot_grid(res_plot_M0Str1_06,res_plot_M1Str1_06, nrow=2))
Residuals_Str2_06<- print(plot_grid(res_plot_M0Str2_06,res_plot_M1Str2_06, nrow=2))
```

### DHARMa tests
```{r}
set.seed(123)
Residuals_test <- data.frame(
  Model = character(),
  Uniformity = character(),
  Uniformity_p = numeric(),
  Dispersion = character(),
  Dispersion_p = numeric(),
  Quantile = character(),
  Quantile_p = numeric(),
  ZeroInfl = character(),
  ZeroInfl_p = numeric(),
  SpatialAut = character(),
  SpatialAut_p = numeric()
)

resid_Chek <- function(model, data, a = 0.05, model_name = NULL) {
  
df_name <- deparse(substitute(data))
model_name <- paste0("model_M1", df_name)
  
# Clean data
data <- data %>% filter(!is.na(Abundance), !is.na(w_sn))

# Simulate and calculate residuals
s_fit <- simulate(model, nsim = 500, type = "mle-mvn")
r_fit <- dharma_residuals(s_fit, model, return_DHARMa = TRUE)
  
# Run DHARMa tests
res_tests <- DHARMa::testResiduals(r_fit, plot = TRUE) 
q_test   <- DHARMa::testQuantiles(r_fit)
zi_test  <- DHARMa::testZeroInflation(r_fit)

# For the spatial autocorrelation 
simulationOutputS <- recalculateResiduals(
    r_fit,
    group = paste(data$X, data$Y),
    aggregateBy = mean)
  
sp_test <- DHARMa::testSpatialAutocorrelation(
   simulationOutputS,
   x = unique(data$X),
   y = unique(data$Y)
 )
  
# Decide pass/fail
 un_re  <- ifelse(res_tests$uniformity$p.value < a, "Fail", "Pass")
 dis_re <- ifelse(res_tests$dispersion$p.value < a, "Fail", "Pass")
 q_res  <- ifelse(q_test$p.value < a, "Fail", "Pass")
 zi_res <- ifelse(zi_test$p.value < a, "Fail", "Pass")
 sp_res <- ifelse(sp_test$p.value < a, "Fail", "Pass")

# Add results to global table
new_row <- data.frame(
  Model = paste(model_name),
  Uniformity = un_re,
  Uniformity_p = round(res_tests$uniformity$p.value, 4),
  Dispersion = dis_re,
  Dispersion_p = round(res_tests$dispersion$p.value, 4),
  Quantile = q_res,
  Quantile_p = round(q_test$p.value, 4),
  ZeroInfl = zi_res,
  ZeroInfl_p = round(zi_test$p.value, 4),
  SpatialAuto = sp_res,
  SpatialAuto_p = round(sp_test$p.value, 4)
  )
  
assign("Residuals_test", rbind(get("Residuals_test", envir = .GlobalEnv), new_row), envir =.GlobalEnv)
}

resid_Chek(model_M1Str1_06, Str1_06)
resid_Chek(model_M1Str2_06, Str2_06)

gt(Residuals_test, caption = "Residuals Test")
Residuals_test_M1 <- Residuals_test
Residuals_test_M1

# Do a summary 
# Select only the test columns
test_cols <- c("Uniformity", "Dispersion", "Quantile", "ZeroInfl", "SpatialAuto")

# Function to count Pass
count_pass <- function(x) sum(x == "Pass")

# Create summary
summary_table <- data.frame(
  Test = test_cols,
  Pass = sapply(Residuals_test[test_cols], count_pass),
  Total = nrow(Residuals_test)
)

# Add Pass Rate as percentage
summary_table$PassRate <- round(summary_table$Pass / summary_table$Total * 100, 1)

# View the summary table
print(summary_table)
Residuals_test_M1_sum <- summary_table
```


# Model 2: GLM with tweedie without spatial autocorralation
Let's move to GLM Tweedie

### Write M2 function
```{r}
fit_glm_TW_M2 <- function(data,plot_basic) {

# Get data frame name
df_name <- deparse(substitute(data))
model_name <- paste0("model_M2", df_name)
Survey <- as.character(data$Survey[1])
dispersion_name  <- paste0("dispersion_M2_", df_name)

data <- data %>% filter(!is.na(Abundance))
data <- data %>% filter(!is.na(w_sn))
Year <- as.character(data$Year[1])

# Build dynamic object names
  plot_name  <- paste0("plot_M2_", df_name)
  plot_log_name  <- paste0("plot_log_M2_", df_name)

  P_est_name <- paste0("P_est_M2_", df_name)
  z_est_name <- paste0("z_est_M2_", df_name)
  
  P_se_name  <- paste0("P_se_M2_", df_name)
  z_se_name  <- paste0("z_se_M2_", df_name)
  
  P_ci_name  <- paste0("P_ci_M2_", df_name)
  z_ci_name  <- paste0("z_ci_M2_", df_name)
  
# Fit model
model2 <- sdmTMB(
  Abundance ~ 1 + (Age),
  data = data,
  family = tweedie(link = "log"),
  weights = data$w_sn,
  spatial = "off")  

assign(model_name, model2, envir = .GlobalEnv)
 
#Summary/diagnostics
summary(model2)
sanity(model2)
  
params <-tidy(model2, conf.int = TRUE)
tidy(model2, effects = "ran_par", conf.int = TRUE)

# Extract params
 P_est <- exp(params$estimate[1]) 
 z_est <- -(params$estimate[2]) 
  
# SEs
  P_se_log <-  params$std.error[1]
  z_se <- params$std.error[2]
  
# CIs form tidy
  P_ci <- NA
  z_ci <-  c(-params$conf.high[2], -params$conf.low[2])
  
# Compute transformed estimates
 #LogP0 
CV_P <- sqrt(exp(P_se_log^2) - 1)
P_se <- P_est * CV_P
P_ci_lower <- exp(params$conf.low[1]) 
P_ci_upper <- exp(params$conf.high[1])

outablePo <- data.frame(
   dataset       =  Survey,
   parameter     = c("P"),
   estimate_log  = params$estimate[1], 
   estimate      = c(P_est),       # Unlog
   se            = c(P_se) ,       #se of the estimated model,
   ci_lower      = c(P_ci_lower),
   ci_upper      = c(P_ci_upper),
   CV            = c(CV_P),        
   model         = "M2",
   stringsAsFactors = FALSE) 

outableZ <- data.frame(
    dataset   = c(Survey),
    parameter = "z",
    estimate  = c(z_est),
    se        = c(z_se),
    ci_lower  = c(z_ci[1]), #It haves been reversed because z insted of -z is modeld
    ci_upper  = c(z_ci[2]),
    CV        = c(z_se/z_est),
    model     = "M2",
    stringsAsFactors = FALSE)


dispersion <-model2$sd_report$value["phi"]%>% as.numeric()
assign(dispersion_name, dispersion, envir = .GlobalEnv)

#Generate predictions for plotting
  x_new <- seq(min(data$Age, na.rm = TRUE),
               max(data$Age, na.rm = TRUE),
               length.out = 100)
  y_pred <- predict(model2, newdata = data.frame(Age = x_new),se_fit = TRUE, re_form = NA) # unlog the prediction
  df_pred <- data.frame(x = x_new, y = exp(y_pred$est),
                        ymin = exp((y_pred$est - 1.96 * (y_pred$est_se))),
                        ymax = exp((y_pred$est + 1.96 * (y_pred$est_se)))) 
  
# Plot
  plot <- 
     plot_basic +
   geom_line(data = df_pred, aes(x, y), color = brewer.pal(7, "Dark2")[3], size = 1) +
   labs(
      title = "",
      subtitle = paste(df_name),
      x = "Age",
      y = "Abundance"
    ) +
    theme_bw() +
    theme(panel.grid.major = element_line(color = "white"))
  
assign(plot_name, plot, envir = .GlobalEnv) 

return(list(outablePo= outablePo, outableZ=outableZ))
}
```

### Fit M2
```{r}
results_all_M2 <- bind_rows(
  fit_glm_TW_M2(Str1_06,plot_M1_Str1_06),     
  fit_glm_TW_M2(Str2_06,plot_M1_Str2_06))
```

### Get P,z results and estimate Ptot
```{r}
outablePo_temp_Str1<- results_all_M2$outablePo %>% filter(grepl("^Str1_",dataset)) %>% mutate(across(where(is.numeric), ~ round(.x, 3)))
outablePo_temp_Str1 <- outablePo_temp_Str1 %>% select (-estimate_log)
Str1_resultsPo <- rbind(Str1_resultsPo,outablePo_temp_Str1)
gt(Str1_resultsPo)

outablePo_temp_Str2<- results_all_M2$outablePo %>% filter(grepl("^Str2",dataset)) %>% mutate(across(where(is.numeric), ~ round(.x, 3)))
outablePo_temp_Str2 <- outablePo_temp_Str2 %>% select (-estimate_log)
Str2_resultsPo <- rbind(Str2_resultsPo,outablePo_temp_Str2)
gt(Str2_resultsPo)

outableZ_temp_Str1 <- results_all_M2$outableZ %>% filter(grepl("^Str1_",dataset)) %>% mutate(across(where(is.numeric), ~ round(.x, 3)))
Str1_resultsZ <- rbind(Str1_resultsZ,outableZ_temp_Str1)
gt(Str1_resultsZ)

outableZ_temp_Str2 <- results_all_M2$outableZ %>% filter(grepl("^Str2_",dataset)) %>% mutate(across(where(is.numeric), ~ round(.x, 3)))
Str2_resultsZ <- rbind(Str2_resultsZ,outableZ_temp_Str2)
gt(Str2_resultsZ)

Str1_resultsPtot_temp <- tibble(
  dataset   = outablePo_temp_Str1$dataset,
  parameter = "Ptot",
  estimate  = outablePo_temp_Str1$estimate * area_Str1,
  se        = NA,
  ci_lower  = outablePo_temp_Str1$ci_lower * area_Str1,
  ci_upper  = outablePo_temp_Str1$ci_upper * area_Str1,
  CV        = outablePo_temp_Str1$CV,
  model     = outablePo_temp_Str1$model
) %>%
  mutate(se = estimate * CV)

Str1_resultsPtot <- rbind(Str1_resultsPtot,Str1_resultsPtot_temp)

Str2_resultsPtot_temp <- tibble(
  dataset   = outablePo_temp_Str2$dataset,
  parameter = "Ptot",
  estimate  = outablePo_temp_Str2$estimate * area_Str2,
  se        = NA,
  ci_lower  = outablePo_temp_Str2$ci_lower * area_Str2,
  ci_upper  = outablePo_temp_Str2$ci_upper * area_Str2,
  CV        = outablePo_temp_Str2$CV,
  model     = outablePo_temp_Str2$model
) %>%
  mutate(se = estimate * CV)

Str2_resultsPtot <- rbind(Str2_resultsPtot,Str2_resultsPtot_temp)
```

### Summary plots
```{r}
# Eastern stratum Po plot
plot_Po_Str1 <- ggplot() +
  geom_point(data = Str1_resultsPo, 
             aes(y = estimate, x = dataset, col = model),
             size = 2, position = position_dodge(width = 0.3)) + 
  geom_errorbar(data = Str1_resultsPo, 
                aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model),
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +  
  theme_bw() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[0] ~ (eggs ~ m^-2 ~ day^-1))) + xlab("") +
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))


# Western stratum Po plot
plot_Po_Str2 <-ggplot() +
  geom_point(data = Str2_resultsPo, 
             aes(y = estimate, x = dataset, col = model),
             size = 2, position = position_dodge(width = 0.3)) + 
  geom_errorbar(data = Str2_resultsPo, 
                aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model), 
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +  
  theme_bw() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[0] ~ (eggs ~ m^-2 ~ day^-1))) + xlab("") +
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))

# Eastern stratum Z plot
plot_z_Str1 <-
ggplot() +
  geom_point(data = Str1_resultsZ, 
             aes(y = estimate, x = dataset, col = model),
             size = 2, position = position_dodge(width = 0.3)) + 
  geom_errorbar(data = Str1_resultsZ, 
                aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model), 
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +  
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(Z ~ (days^-1))) + xlab("") +
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))


# Western stratum Z plot
plot_z_Str2 <-
ggplot() +
  geom_point(data = Str2_resultsZ,
             aes(y = estimate, x = dataset, col = model),
             size = 2, position = position_dodge(width = 0.3)) + 
  geom_errorbar(data = Str2_resultsZ, 
                aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model), 
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +  
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(Z ~ (day^-1))) + xlab("") +
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))

# Eastern stratum Ptot plot
plot_Ptot_Str1 <-ggplot() +
  geom_point(data = Str1_resultsPtot, 
             aes(y = estimate / 10^11, x = dataset, col = model), 
             size = 2, position =position_dodge(width = 0.3)) + 
  geom_errorbar(data = Str1_resultsPtot, 
                aes(x = dataset, ymin = ci_lower / 10^11, ymax = ci_upper / 10^11, color = model),
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +   
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[tot] ~ (eggs ~ day^-1) ~ x10^11)) + xlab("") +
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))

# Western stratum Ptot plot
plot_Ptot_Str2 <-ggplot() +
  geom_point(data = Str2_resultsPtot, 
             aes(y = estimate / 10^11, x = dataset, col = model),
             size = 2, position = position_dodge(width = 0.3)) + 
  geom_errorbar(data = Str2_resultsPtot, 
                aes(x = dataset, ymin = ci_lower / 10^11, ymax = ci_upper / 10^11, color = model),
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +   
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[tot] ~ (eggs ~ day^-1) ~ x10^11)) + xlab("") +
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))

Summary_Plots <-  
  print(plot_grid(
    plot_Po_Str1, plot_Po_Str2,
    plot_z_Str1, plot_z_Str2, 
    plot_Ptot_Str1, plot_Ptot_Str2,
    ncol = 2))
```

### Visualize Fitts 
```{r}
#Plots of fitted models
Ploted_fits_Str1<- print(plot_grid(plot_M2_Str1_06))
Ploted_fits_Str2<- print(plot_grid(plot_M2_Str2_06))
```

### QQ and Residual
```{r}
Residuals_plots<- function(data,model,model_name){
    df_name <- deparse(substitute(data))
    res_plot_name  <- paste0("res_plot_",model_name,df_name)
    res_qq_name  <- paste0("res_qq_",model_name,df_name)
    data <- data %>% filter(!is.na(Abundance), !is.na(w_sn))
   year <- sub(".*_(\\d{2}).*", "\\1", df_name) ;year <- paste0("20", year)
  
 # Determine stratum
 stratum <- case_when(
  grepl("^Str1", df_name) ~ "Eastern stratum",
  grepl("^Str2", df_name) ~ "Western stratum",
  TRUE ~ ""
)
 
df_res <- data.frame(
  Age = data$Age,
  Predicted = exp((predict(model, newdata = data.frame(Age = data$Age)))$est), # Unlog
  Observed= data$Abundance) %>% mutate (Residual= residuals(model,type ="mle-mvn"))

#Pearson Cor  
r_val <- cor(df_res$Predicted, df_res$Observed, use = "complete.obs",method = "pearson")

# Residual vs Fitted plot
p_obs_pred <- ggplot(df_res, aes(x = Predicted, y = Observed)) +
  geom_point(alpha = 0.6,color=brewer.pal(7, "Dark2")[3]) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
 
  labs(
    x = "",
    y = "Observed",
    subtitle = paste(model_name) )+
annotate(
  "text",
  x = -Inf,
  y = Inf,
  label = sprintf("r = %.3f", r_val),
  hjust = -0.1,
  vjust = 1.1,
  size =4, color="red") +
  theme_bw() +
  theme(plot.margin = margin(t=0,r = 1,b =0,l = 1))

p_res_fit <- ggplot(df_res, aes(x = Predicted, y = Residual)) +
    geom_point(alpha = 0.6,color=brewer.pal(7, "Dark2")[3]) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    theme_bw() +
    labs(
      x = "",
      y = "Residuals", 
      subtitle="") +
     theme_bw() +
     theme(plot.margin = margin(t=0,r = 1,b =0,l = 1))


p_qq <- ggplot(df_res, aes(sample = Residual)) +
    stat_qq(color=brewer.pal(7, "Dark2")[3]) +
    stat_qq_line(linetype = "dashed",color = "red") +
    theme_bw() +
    labs(
      x = "",
      y = "Sample Q",
     subtitle="") +
    theme(plot.margin = margin(t=0,r = 1,b =0,l = 1))

res_plots<- plot_grid(p_obs_pred,p_res_fit,p_qq, ncol=3)
assign(res_plot_name, res_plots, envir = .GlobalEnv)  
assign(res_qq_name, p_qq, envir = .GlobalEnv)  
}

Residuals_plots(Str1_06,model_M2Str1_06, "M2") ; res_plot_M2Str1_06 ;res_qq_M2Str1_06
Residuals_plots(Str2_06,model_M2Str2_06, "M2") ; res_plot_M2Str2_06 ;res_qq_M2Str2_06

Residuals_Str1_06<- print(plot_grid(res_plot_M0Str1_06,res_plot_M1Str1_06,res_plot_M2Str1_06, nrow=3))
Residuals_Str2_06<- print(plot_grid(res_plot_M0Str2_06,res_plot_M1Str2_06,res_plot_M2Str2_06, nrow=3))
```

### DHAMR residuals 
```{r}
set.seed(123)
Residuals_test <- data.frame(
    Model = character(),
    Uniformity = character(),
    Uniformity_p = numeric(),
    Dispersion = character(),
    Dispersion_p= numeric(),
    Quantile   = character(),
    Quantile_p = numeric(),
    ZeroInfl   = character(),
    ZeroInfl_p = numeric(),
    SpatialAut = character(),
    SpatialAut_p= numeric())

resid_Chek <- function(model, data, a = 0.05, model_name = NULL) {
  
  df_name <- deparse(substitute(data))
  model_name <- paste0("model_M2", df_name)
  
  # Clean data
  data <- data %>% filter(!is.na(Abundance), !is.na(w_sn))

  # Simulate and calculate residuals
  s_fit <- simulate(model, nsim = 500, type = "mle-mvn")
  r_fit <- dharma_residuals(s_fit, model, return_DHARMa = TRUE)
  
  # Run DHARMa tests
  res_tests<-DHARMa::testResiduals(r_fit,plot=T)  
  q_test   <- DHARMa::testQuantiles(r_fit) #
  zi_test  <- DHARMa::testZeroInflation(r_fit)

# For the spatial autocorelation 
  simulationOutputS <- recalculateResiduals(
    r_fit,
    group = paste(data$X, data$Y),
    aggregateBy = mean  )
  
  sp_test <- DHARMa::testSpatialAutocorrelation(
    simulationOutputS,
    x = unique(data$X),
    y = unique(data$Y)
  )
  
  # Decide pass/fail
  un_re  <- ifelse(res_tests$uniformity$p.value  < a, "Fail", "Pass")
  dis_re  <- ifelse(res_tests$dispersion$p.value  < a, "Fail", "Pass")
  q_res  <- ifelse(q_test$p.value  < a, "Fail", "Pass")
  zi_res <- ifelse(zi_test$p.value < a, "Fail", "Pass")
  sp_res <- ifelse(sp_test$p.value < a, "Fail", "Pass")
  
  # Add results to global table
  new_row <- data.frame(
    Model = paste(model_name),
    Uniformity = un_re,
    Uniformity_p = round(res_tests$uniformity$p.value,4),
    Dispersion = dis_re,
    Dispersion_p = round(res_tests$dispersion$p.value,4),
    Quantile = q_res,
    Quantile_p= round(q_test$p.value,4),
    ZeroInfl = zi_res,
    ZeroInfl_p= round(zi_test$p.value,4),
    SpatialAuto = sp_res,
    SpatialAuto_p= round(sp_test$p.value,4))
  
assign("Residuals_test", rbind(get("Residuals_test", envir = .GlobalEnv), new_row), envir = .GlobalEnv)
}

 resid_Chek(model_M2Str1_06,Str1_06)
 resid_Chek(model_M2Str2_06,Str2_06)

gt(Residuals_test,caption = "Residuals Test")
Residuals_test_M2 <- Residuals_test

# Do a summary 
# Select only the test columns
test_cols <- c("Uniformity", "Dispersion", "Quantile", "ZeroInfl", "SpatialAuto")

# Function to count Pass
count_pass <- function(x) sum(x == "Pass")

# Create summary
summary_table <- data.frame(
  Test = test_cols,
  Pass = sapply(Residuals_test[test_cols], count_pass),
  Total = nrow(Residuals_test)
)

# Add Pass Rate as percentage
summary_table$PassRate <- round(summary_table$Pass / summary_table$Total * 100, 1)

# View the summary table
print(summary_table)
Residuals_test_M2_sum <- summary_table
```

# Some statistiscs to compaire models
```{r}
#Eastern stratum datasets
data <- Str1_06 %>%filter(!is.na(Abundance), !is.na(w_sn))
pred_M0 <- predict(model_M0Str1_06, type = "response")
pred_M1 <- predict(model_M1Str1_06, type = "response")
pred_M2 <- predict(model_M2Str1_06, type = "response")
obs <- data$Abundance

rmse_Str1_06_M0 <- sqrt(mean((obs-pred_M0)^2, na.rm = TRUE))
rmse_Str1_06_M1 <- sqrt(mean((obs-pred_M1$est)^2, na.rm = TRUE))
rmse_Str1_06_M2 <- sqrt(mean((obs-pred_M2$est)^2, na.rm = TRUE))
cor_Str1_06_M0  <- cor(obs, pred_M0, use = "complete.obs", method = "pearson")
cor_Str1_06_M1  <- cor(obs, pred_M1$est, use = "complete.obs", method = "pearson")
cor_Str1_06_M2  <- cor(obs, pred_M2$est, use = "complete.obs", method = "pearson")

#Western Stratum datasets
data <- Str2_06 %>%filter(!is.na(Abundance), !is.na(w_sn))
pred_M0 <- predict(model_M0Str2_06, type = "response")
pred_M1 <- predict(model_M1Str2_06, type = "response")
pred_M2 <- predict(model_M2Str2_06, type = "response")
obs <- data$Abundance

rmse_Str2_06_M0 <- sqrt(mean((obs-pred_M0)^2, na.rm = TRUE))
rmse_Str2_06_M1 <- sqrt(mean((obs-pred_M1$est)^2, na.rm = TRUE))
rmse_Str2_06_M2 <- sqrt(mean((obs-pred_M2$est)^2, na.rm = TRUE))
cor_Str2_06_M0  <- cor(obs, pred_M0, use = "complete.obs", method = "pearson")
cor_Str2_06_M1  <- cor(obs, pred_M1$est, use = "complete.obs", method = "pearson")
cor_Str2_06_M2  <- cor(obs, pred_M2$est, use = "complete.obs", method = "pearson")

#Merge
RMSE<- tibble(dataset= c("Str1_06","Str2_06"),
              RMSE_M0= c(rmse_Str1_06_M0,rmse_Str2_06_M0),            
              RMSE_M1= c(rmse_Str1_06_M1,rmse_Str2_06_M2),
              RMSE_M2= c(rmse_Str1_06_M2,rmse_Str2_06_M2),
              COR_M0 = c(cor_Str1_06_M0,cor_Str2_06_M0),
              COR_M1 = c(cor_Str1_06_M1,cor_Str1_06_M1),
              COR_M2 = c(cor_Str1_06_M2,cor_Str1_06_M2))
```

# Constract the Barrier mesh  need to run the Spatial Models
```{r}
#  Create factions to create for the barrier meshes
create_mesh_Str1<- function(data,PA_polygone,edge_est_km)
{
df_name <- deparse(substitute(data))
mesh_name <- paste0("bspde", df_name)

data <- data %>% filter(!is.na(Abundance))
data <- data %>% filter(!is.na(w_sn))
data.eggs <- data

poly.coast <- coast # in WGS84 
stations_sf <- st_as_sf(data.eggs, coords = c("Lon","Lat"), crs = 4326) 
plot(stations_sf$geometry, pch=19, col=2)

#Create a convex hull polygon around all stations 
transect_poly <- st_convex_hull(st_union(stations_sf))

# Buffer around the transect polygon (e.g., 0.1 degrees ~ ~10km depending on latitude)
buff <- st_buffer(transect_poly, dist = 10000) # WGS84 degrees; or use UTM for meters
plot(buff, col = "lightblue", border = "blue");plot(stations_sf$geometry, add = TRUE, pch = 19, col = 2)

output = st_intersection(buff,poly.coast)
plot(output,col=3)
domain = st_difference(transect_poly,st_union(output)) 
domain = st_sf(domain)

domain_utm <- st_transform(domain, "+init=epsg:32635 +units=km")
plot(domain_utm,col=4,axes=T);plot(data.eggs$X, data.eggs$Y, pch="+",add=T)

range_est<- edge_est_km*5 
max.edge = edge_est_km 
bound.outer = range_est*(2/3)

poly.coast <- st_make_valid(poly.coast)
domain_utm <- st_make_valid(domain_utm)

# Build high-resolution mesh
mesh_high = inla.mesh.2d( 
  boundary = domain_utm, 
  loc = cbind(data.eggs$X, data.eggs$Y),
  max.edge = c(1,5)*max.edge,
  cutoff = max.edge/5, 
  offset = c(max.edge, bound.outer)
  )

plot(mesh_high, lwd=0.5, axes = T); points(data.eggs$X, data.eggs$Y, col="red")
mesh_high <- make_mesh(data.eggs, c("X", "Y"), mesh = mesh_high) 
plot(mesh_high, lwd=0.5, axes = T); points(data.eggs$X, data.eggs$Y, col="red")

# Add on the barrier mesh component
poly.coast_utm <- st_transform(poly.coast, "+init=epsg:32635 +units=km")
bspde <- add_barrier_mesh(
  spde_obj = mesh_high,
  barrier_sf = poly.coast_utm,
  range_fraction = 0.1,
  proj_scaling = 1,
  plot = TRUE)
plot(bspde, lwd=0.5); points(data.eggs$X, data.eggs$Y, col="red")
assign(mesh_name, bspde, envir = .GlobalEnv)
}

create_mesh_Str2<- function(data,PA_polygone,edge_est_km)
{
df_name <- deparse(substitute(data))
mesh_name <- paste0("bspde", df_name)

data <- data %>% filter(!is.na(Abundance))
data <- data %>% filter(!is.na(w_sn))
data.eggs <- data

poly.coast <- coast # in WGS84 
stations_sf <- st_as_sf(data.eggs, coords = c("Lon","Lat"), crs = 4326) 
plot(stations_sf$geometry, pch=19, col=2)

#Create a convex hull polygon around all stations 
transect_poly <- st_convex_hull(st_union(stations_sf))

# Buffer around the transect polygon (e.g., 0.1 degrees ~ ~10km depending on latitude)
buff <- st_buffer(transect_poly, dist = 10000) # WGS84 degrees; or use UTM for meters
plot(buff, col = "lightblue", border = "blue");plot(stations_sf$geometry, add = TRUE, pch = 19, col = 2)

output = st_intersection(buff,poly.coast)
plot(output,col=3)
domain = st_difference(transect_poly,st_union(output)) 
domain = st_sf(domain)

domain_utm <- st_transform(domain, "+init=epsg:32635 +units=km")
plot(domain_utm,col=4,axes=T);plot(data.eggs$X, data.eggs$Y, pch="+",add=T)

# Mesh characteristics 
range_est<-  edge_est_km*5 
max.edge =edge_est_km  
bound.outer = range_est*(2/3)

#force geometries to be valid:
poly.coast <- st_make_valid(poly.coast)
domain_utm <- st_make_valid(domain_utm)

# Build high-resolution mesh
mesh_high = inla.mesh.2d( 
  boundary = domain_utm, 
  loc = cbind(data.eggs$X, data.eggs$Y),
  max.edge = c(1,5)*max.edge,
  cutoff = max.edge/5, 
  offset = c(max.edge, bound.outer)
  )

plot(mesh_high, lwd=0.5, axes = T); points(data.eggs$X, data.eggs$Y, col="red")
mesh_high <- make_mesh(data.eggs, c("X", "Y"), mesh = mesh_high) 
plot(mesh_high, lwd=0.5, axes = T); points(data.eggs$X, data.eggs$Y, col="red")

# Add on the barrier mesh component
poly.coast_utm <- st_transform(poly.coast, "+init=epsg:32635 +units=km")
bspde <- add_barrier_mesh(
  spde_obj = mesh_high,
  barrier_sf = poly.coast_utm,
  range_fraction = 0.1,
  proj_scaling = 1,
  plot = TRUE)

plot(bspde, lwd=0.5); points(data.eggs$X, data.eggs$Y, col="red")
assign(mesh_name, bspde, envir = .GlobalEnv)
}

# Crate the meshes using the optimized edge value selacted based on the prior sensitivity analyses
create_mesh_Str1(Str1_06, PA_2006_Str1,edge_est_km=22) 
create_mesh_Str2(Str2_06,PA_2006_Str2,edge_est_km=7)  
``` 

# Model 3: GLMM with NB and spatial autocorralation

### Write M3 function 
```{r}
fit_glm_NBsp_M3 <- function(data,PA_polygone,bspde,plot_basic) {
 
# Get data frame name
df_name <- deparse(substitute(data))
model_name <- paste0("model_M3", df_name)
Survey <- as.character(data$Survey[1])
dispersion_name  <- paste0("dispersion_M3_", df_name)

data <- data %>% filter(!is.na(Abundance))
data <- data %>% filter(!is.na(w_sn))
Year <- as.character(data$Year[1])

# Build dynamic object names
  plot_name  <- paste0("plot_M3_", df_name)
  plot_log_name  <- paste0("plot_log_M3_", df_name)
  omega_name  <- paste0("omegaSD_M3_", df_name)
  range_name  <- paste0("range_M3_", df_name)
  dispersion_name  <- paste0("dispersion_M3_", df_name)

  P_est_name <- paste0("P_est_M3_", df_name)
  z_est_name <- paste0("z_est_M3_", df_name)
  
  P_se_name  <- paste0("P_se_M3_", df_name)
  z_se_name  <- paste0("z_se_M3_", df_name)
  
  P_ci_name  <- paste0("P_ci_M3_", df_name)
  z_ci_name  <- paste0("z_ci_M3_", df_name)
  
# Fit model
model3 <-  sdmTMB(
  Ni ~ 1 + (Age),
  data = data,
  offset= log(data$Ri), # log because we are at the logit 
  family = nbinom2(link = "log"),
  mesh = bspde,
  spatial = "on")  

assign(model_name, model3, envir = .GlobalEnv)

#Summary
summary(model3)
sanity(model3)

params <-tidy(model3, conf.int = TRUE)
tidy(model3, effects = "ran_par", conf.int = TRUE)

#Extract param
 P_est <- exp(params$estimate[1]) 
 z_est <- -(params$estimate[2]) 
  
#SE
  P_se_log <-  params$std.error[1]
  z_se <- params$std.error[2]
  
#CI 
  P_ci <- NA  
  z_ci <-  c(-params$conf.high[2], -params$conf.low[2]) 
  
#Spatial stracture
 rep <- model3$tmb_obj$report()
 sigma_O <- rep$sigma_O [1,1]
 range<-  rep$range[1,1]
 dispersion <-model3$sd_report$value["phi"]%>% as.numeric()

assign(omega_name, sigma_O, envir = .GlobalEnv)
assign(range_name, range, envir = .GlobalEnv)
assign(dispersion_name, dispersion, envir = .GlobalEnv)

# Compute transformed estimates
#LogP0
CV_P <- sqrt(exp(P_se_log^2) - 1)
P_se <- P_est * CV_P
P_ci_lower <- exp(params$conf.low[1]) 
P_ci_upper <- exp(params$conf.high[1])

outablePo <- data.frame(
    dataset       =  Survey,
    parameter     = c("P"),
    estimate_log  = params$estimate[1], 
    estimate      = c(P_est),       # Unlog
    se            = c(P_se) ,       #se of the estimated model,
    ci_lower      = c(P_ci_lower),
    ci_upper      = c(P_ci_upper),
    CV            = c(CV_P),        
    model         = "M3",
    stringsAsFactors = FALSE) 

  outableZ <- data.frame(
    dataset   = c(Survey),
    parameter = "z",
    estimate  = c(z_est),
    se        = c(z_se),
    ci_lower  = c(z_ci[1]), #It haves been reversed because z insted of -z is modeld
    ci_upper  = c(z_ci[2]),
    CV        = c(z_se/z_est),
    model     = "M3",
    stringsAsFactors = FALSE)

#Predict in data points 
# Back predict on the data:  for the  omega -and the  the plots 
predic.grid1 <- data %>% filter(!is.na(Stage))  %>% filter(!is.na(Age))  
predict1 <- predict(model3,newdata=predic.grid1,re_form = NA) #Turn of the spatial filed in the prediction

model_prediction1 <- paste0("model_M3", df_name, "_pred1")
assign(model_prediction1, predict1, envir = .GlobalEnv)

# Plot
  plot <- 
     plot_basic +
   geom_line(data = predict1, aes(x=Age, y= exp(est)), color = brewer.pal(7, "Dark2")[4], linewidth = 1) +
   labs(
      title = "",
      subtitle = paste(df_name),#, ": P=", round(P_est, 3), ", z=", round(z_est, 3)),
      x = "Age",
      y = "Abundance" ) +
    theme_bw() +
    theme(panel.grid.major = element_line(color = "white"))
  
assign(plot_name, plot, envir = .GlobalEnv) 

#For the Ptot Prediction -> index standardization
#1. Create the extrapolation grid (PA- division)
grid_1km <- st_make_grid(
  PA_polygone,
  cellsize = 1,        # 1 unit = 1 km
  square = TRUE,
  what = "polygons")

grid_clipped <- st_intersection(grid_1km, PA_polygone)  # Do the grid only in the Positive area 

plot(grid_clipped)
centroids <- st_centroid(grid_clipped)
centroids_df <- st_coordinates(centroids) %>%
  as.data.frame() 

centroids_m <- centroids %>% st_sf()  # ensure sf
st_geometry(centroids_m) <- st_geometry(centroids_m) * 1000  # km -> m
st_crs(centroids_m) <- 32635
centroids_wgs84 <- st_transform(centroids_m, crs = 4326)

lonlat_df <- st_coordinates(centroids_wgs84) %>%
  as.data.frame() %>% dplyr::rename(Lon = X, Lat = Y)

centroids_df <- bind_cols(centroids_df, lonlat_df)

#This is the new prediction grid that will force the prediction index standardization 
#log(Pij)=log(Po) + ω +  (z*t) , but z*t=0 so --> log(Pij)=log(Po) + ω (Prediction is now Ptot)
predic.grid2<- centroids_df %>% mutate(Age=0)

#Predict
predict2 <- predict(model3,newdata=predic.grid2,return_tmb_object = TRUE) 
model_prediction2 <- paste0("model_M3", df_name, "_pred2")
assign(model_prediction2, predict2, envir = .GlobalEnv)
 
#3.Get the index (Ptot)
index <- get_index(predict2, area =10^6, bias_correct = TRUE)  # 10^6 for making kM1--> M1
index <- index %>% mutate(cv = sqrt(exp(se^2) - 1))   #Since se refer to the log estimate

outablelPtot <- data.frame(
  dataset       =  Survey,
  parameter = "Ptot",
  estimate  = index$est, #estimate in the original scale
  se        =  index$cv*index$est,  # re-estimate it because it was in log(abundance)
  ci_lower  =  index$lwr,
  ci_upper  =  index$upr,
  CV        = index$cv,   # re-estimate it because it was in log(abundance))
  model         = "M3",
  stringsAsFactors = FALSE) 

# Output tables
return(list(outablePo= outablePo, outableZ=outableZ,outablelPtot=outablelPtot))
}
```

### Fit M3
```{r}
results_all_M3 <- bind_rows(
fit_glm_NBsp_M3(Str1_06,PA_2006_Str1,bspdeStr1_06,plot_M2_Str1_06),
fit_glm_NBsp_M3(Str2_06,PA_2006_Str2,bspdeStr2_06,plot_M2_Str2_06))
```

### Get P,z results and estimate Ptot
```{r}
outablePo_temp_Str1<- results_all_M3$outablePo %>% filter(grepl("^Str1_",dataset)) %>% mutate(across(where(is.numeric), ~ round(.x, 3)))
outablePo_temp_Str1 <- outablePo_temp_Str1 %>% select (-estimate_log)
Str1_resultsPo <- rbind(Str1_resultsPo,outablePo_temp_Str1)
gt(Str1_resultsPo)

outablePo_temp_Str2<- results_all_M3$outablePo %>% filter(grepl("^Str2",dataset)) %>% mutate(across(where(is.numeric), ~ round(.x, 3)))
outablePo_temp_Str2 <- outablePo_temp_Str2 %>% select (-estimate_log)
Str2_resultsPo <- rbind(Str2_resultsPo,outablePo_temp_Str2)
gt(Str2_resultsPo)

outableZ_temp_Str1 <- results_all_M3$outableZ %>% filter(grepl("^Str1_",dataset)) %>% mutate(across(where(is.numeric), ~ round(.x, 3)))
Str1_resultsZ <- rbind(Str1_resultsZ,outableZ_temp_Str1)
gt(Str1_resultsZ)

outableZ_temp_Str2 <- results_all_M3$outableZ %>% filter(grepl("^Str2_",dataset)) %>% mutate(across(where(is.numeric), ~ round(.x, 3)))
Str2_resultsZ <- rbind(Str2_resultsZ,outableZ_temp_Str2)
gt(Str2_resultsZ)

#Plot will be prediction output 
Str1_resultsPtot_temp <-results_all_M3$outablelPtot %>% filter(grepl("^Str1_",dataset)) 

Str1_resultsPtot <- rbind(Str1_resultsPtot,Str1_resultsPtot_temp)
gt(Str1_resultsPtot)

Str2_resultsPtot_temp <-results_all_M3$outablelPtot %>% filter(grepl("^Str2_",dataset)) 

Str2_resultsPtot <- rbind(Str2_resultsPtot,Str2_resultsPtot_temp)
gt(Str2_resultsPtot)
```

### Summary plots
```{r}
# Eastern stratum Po plot
plot_Po_Str1 <- ggplot() +
  geom_point(data = Str1_resultsPo, 
             aes(y = estimate, x = dataset, col = model),
             size = 2, position = position_dodge(width = 0.3)) + 
  geom_errorbar(data = Str1_resultsPo, 
                aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model),
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +  
  theme_bw() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[0] ~ (eggs ~ m^-2 ~ day^-1))) + xlab("") +
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))


# Western stratum Po plot
plot_Po_Str2 <-ggplot() +
  geom_point(data = Str2_resultsPo, 
             aes(y = estimate, x = dataset, col = model),
             size = 2, position = position_dodge(width = 0.3)) + 
  geom_errorbar(data = Str2_resultsPo, 
                aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model), 
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +  
  theme_bw() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[0] ~ (eggs ~ m^-2 ~ day^-1))) + xlab("") +
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))

# Eastern stratum Z plot
plot_z_Str1 <-
ggplot() +
  geom_point(data = Str1_resultsZ, 
             aes(y = estimate, x = dataset, col = model),
             size = 2, position = position_dodge(width = 0.3)) + 
  geom_errorbar(data = Str1_resultsZ, 
                aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model), 
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +  
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(Z ~ (days^-1))) + xlab("") +
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))


# Western stratum Z plot
plot_z_Str2 <-
ggplot() +
  geom_point(data = Str2_resultsZ,
             aes(y = estimate, x = dataset, col = model),
             size = 2, position = position_dodge(width = 0.3)) + 
  geom_errorbar(data = Str2_resultsZ, 
                aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model), 
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +  
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(Z ~ (day^-1))) + xlab("") +
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))

# Eastern stratum Ptot plot
plot_Ptot_Str1 <-ggplot() +
  geom_point(data = Str1_resultsPtot, 
             aes(y = estimate / 10^11, x = dataset, col = model), 
             size = 2, position =position_dodge(width = 0.3)) + 
  geom_errorbar(data = Str1_resultsPtot, 
                aes(x = dataset, ymin = ci_lower / 10^11, ymax = ci_upper / 10^11, color = model),
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +   
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[tot] ~ (eggs ~ day^-1) ~ x10^11)) + xlab("") +
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))

# Western stratum Ptot plot
plot_Ptot_Str2 <-ggplot() +
  geom_point(data = Str2_resultsPtot, 
             aes(y = estimate / 10^11, x = dataset, col = model),
             size = 2, position = position_dodge(width = 0.3)) + 
  geom_errorbar(data = Str2_resultsPtot, 
                aes(x = dataset, ymin = ci_lower / 10^11, ymax = ci_upper / 10^11, color = model),
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +   
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[tot] ~ (eggs ~ day^-1) ~ x10^11)) + xlab("") +
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))

Summary_Plots <-  
  print(plot_grid(
    plot_Po_Str1, plot_Po_Str2,
    plot_z_Str1, plot_z_Str2, 
    plot_Ptot_Str1, plot_Ptot_Str2,
    ncol = 2))
```

### Visualize Fits 
```{r}
Ploted_fits_Str1<- print(plot_grid(plot_M3_Str1_06))
Ploted_fits_Str2<- print(plot_grid(plot_M3_Str2_06))
```

### QQ and Residuals 
```{r}
Residuals_plots<- function(data,model,model_name){
    df_name <- deparse(substitute(data))
    res_plot_name  <- paste0("res_plot_",model_name,df_name)
    res_qq_name  <- paste0("res_qq_",model_name,df_name)
    data <- data %>% filter(!is.na(Abundance), !is.na(w_sn))
    year <- sub(".*_(\\d{2}).*", "\\1", df_name) ;year <- paste0("20", year)
   
 # Determine stratum
 stratum <- case_when(
  grepl("^Str1", df_name) ~ "Eastern stratum",
  grepl("^Str2", df_name) ~ "Western stratum",
  TRUE ~ ""
)
# Generate predictions on the original Age values
df_res <- data.frame(
  Age = data$Age,
  Predicted = exp((predict(model, newdata = data.frame(Age = data$Age,X=data$X,Y=data$Y)))$est), 
  Observed= data$Abundance) %>% mutate (Residual= residuals(model,type ="mle-mvn"))

#Pearson Cor coef 
r_val <- cor(df_res$Predicted, df_res$Observed, use = "complete.obs",method = "pearson")

# Residual vs Fitted plot
p_obs_pred <- ggplot(df_res, aes(x = Predicted, y = Observed)) +
  geom_point(alpha = 0.6,color=brewer.pal(7, "Dark2")[4]) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(
    x = "",
    y = "Observed",
    subtitle = paste(model_name) )+
  annotate(
   "text",
   x = -Inf,
   y = Inf,
   label = sprintf("r = %.3f", r_val),
   hjust = -0.1,
   vjust = 1.1,
   size =4, color="red")  + 
 theme_bw() +
 theme(plot.margin = margin(t=0,r=1,b=0,l=1))
   
p_res_fit <- ggplot(df_res, aes(x = Predicted, y = Residual)) +
    geom_point(alpha = 0.6,color=brewer.pal(7, "Dark2")[4]) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    theme_bw() +
    labs(
      y = "Residuals",
      x = "",
      subtitle="")+ 
   theme_bw() +
   theme(plot.margin = margin(t =0,r = 1,b =0,l = 1))

p_qq <- ggplot(df_res, aes(sample = Residual)) +
    stat_qq(color=brewer.pal(7, "Dark2")[4]) +
    stat_qq_line(linetype = "dashed",color = "red") +
    theme_bw() +
    labs(
      x = "",
      y = "Sample Q",
      subtitle="") + 
     theme(plot.margin = margin(t=0,r = 1,b =0,l = 1))

res_plots<- plot_grid(p_obs_pred,p_res_fit,p_qq, ncol=3)

assign(res_plot_name, res_plots, envir = .GlobalEnv) 
assign(res_qq_name, res_plots, envir = .GlobalEnv)  
}

Residuals_plots(Str1_06,model_M3Str1_06, "M3") ; res_plot_M3Str1_06 ;res_qq_M3Str1_06
Residuals_plots(Str2_06,model_M3Str2_06, "M3") ; res_plot_M3Str2_06 ;res_qq_M3Str2_06

Residuals_Str1_06<- print(plot_grid(res_plot_M0Str1_06,res_plot_M1Str1_06,res_plot_M2Str1_06,res_plot_M3Str1_06,nrow=4))
Residuals_Str2_06<- print(plot_grid(res_plot_M0Str2_06,res_plot_M1Str2_06,res_plot_M2Str2_06,res_plot_M3Str2_06, nrow=4))
```

### DHARMa residuals 
```{r}
set.seed(123) # So you get the same output in any simulation 

Residuals_test <- data.frame(
    Model = character(),
    Uniformity = character(),
    Uniformity_p = numeric(),
    Dispersion = character(),
    Dispersion_p= numeric(),
    Quantile   = character(),
    Quantile_p = numeric(),
    ZeroInfl   = character(),
    ZeroInfl_p = numeric(),
    SpatialAut = character(),
    SpatialAut_p= numeric())

resid_Chek <- function(model, data, a = 0.05, model_name = NULL) {
  
df_name <- deparse(substitute(data))
  model_name <- paste0("model_M3", df_name)
  
# Clean data
  data <- data %>% filter(!is.na(Abundance), !is.na(w_sn))

# Simulate and calculate residuals
  s_fit <- simulate(model, nsim = 500, type = "mle-mvn")
  r_fit <- dharma_residuals(s_fit, model, return_DHARMa = TRUE)
  
# Run DHARMa tests
  res_tests<-DHARMa::testResiduals(r_fit,plot=T)  
  q_test   <- DHARMa::testQuantiles(r_fit) # same as plotResiduals quantile residuals versus (ranked) fitted values. Check for homogeneity of variance-
  zi_test  <- DHARMa::testZeroInflation(r_fit)

# For the spatial autocorelation 
simulationOutputS <- recalculateResiduals(
    r_fit,
    group = paste(data$X, data$Y),
    aggregateBy = mean)
  
sp_test <- DHARMa::testSpatialAutocorrelation(
    simulationOutputS,
    x = unique(data$X),
    y = unique(data$Y))
  
# Decide pass/fail
un_re  <- ifelse(res_tests$uniformity$p.value  < a, "Fail", "Pass")
dis_re  <- ifelse(res_tests$dispersion$p.value  < a, "Fail", "Pass")
q_res  <- ifelse(q_test$p.value  < a, "Fail", "Pass")
zi_res <- ifelse(zi_test$p.value < a, "Fail", "Pass")
sp_res <- ifelse(sp_test$p.value < a, "Fail", "Pass")
  
# Add results to global table
new_row <- data.frame(
  Model = paste(model_name),
  Uniformity = un_re,
  Uniformity_p = round(res_tests$uniformity$p.value,4),
  Dispersion = dis_re,
  Dispersion_p = round(res_tests$dispersion$p.value,4),
  Quantile = q_res,
  Quantile_p= round(q_test$p.value,4),
  ZeroInfl = zi_res,
  ZeroInfl_p= round(zi_test$p.value,4),
  SpatialAuto = sp_res,
  SpatialAuto_p= round(sp_test$p.value,4) )
  
assign("Residuals_test", rbind(get("Residuals_test", envir = .GlobalEnv), new_row), envir = .GlobalEnv)

}

 resid_Chek(model_M3Str1_06,Str1_06)
 resid_Chek(model_M3Str2_06,Str2_06)

gt(Residuals_test,caption = "Residuals Test")
Residuals_test_M3 <- Residuals_test

# Do a summary 
# Select only the test columns
test_cols <- c("Uniformity", "Dispersion", "Quantile", "ZeroInfl", "SpatialAuto")

# Function to count Pass
count_pass <- function(x) sum(x == "Pass")

# Create summary
summary_table <- data.frame(
  Test = test_cols,
  Pass = sapply(Residuals_test[test_cols], count_pass),
  Total = nrow(Residuals_test))

# Add Pass Rate as percentage
summary_table$PassRate <- round(summary_table$Pass / summary_table$Total * 100, 1)

# View the summary table
print(summary_table)
Residuals_test_M3_sum <- summary_table
```

### Omegas
```{r}
#plot all the omegas 
all_omega_Str1 <- c(model_M3Str1_06_pred2$data$omega_s)
color_limits_Str1 <- c(min(all_omega_Str1, na.rm = TRUE),max(all_omega_Str1, na.rm = TRUE)) 

all_omega_Str2 <- c(model_M3Str2_06_pred2$data$omega_s)
color_limits_Str2 <- c(min(all_omega_Str2, na.rm = TRUE),max(all_omega_Str2, na.rm = TRUE)) 

tile_size <- 0.015
tile_width <- 1 / (111 * cos(25 * pi/180))  # ~0.0102°
tile_height <- 1 / 111  # ~0.0090°
tile_height <- tile_height * 1.15
tile_width <- tile_width * 1.2

plot_O_pub_Str1 <- function(prediction, m_code, legend = TRUE) {
p<- ggplot() +
   # Coastline
  geom_tile(
    data = prediction,
    aes(x = Lon, y = Lat, fill = omega_s),
    width = tile_width,  # adjust to your grid spacing
    height = tile_height
  ) +
   scale_fill_gradient2(name = expression(omega[s]),
     low = "red", mid = "white", high = "blue",
     midpoint = 0,
     limits = color_limits_Str1)+
    labs(x = "Longitude", y = "Latitude", title = m_code) +
    theme(panel.background = element_rect(fill = "gray85", color = NA))+
    geom_sf(data = coast, fill = "gray20", col = "gray50") +
    coord_sf(xlim = c(23.5, 26), ylim = c(39.3, 41)) +
    scale_x_continuous(breaks = seq(23.5, 26.5, by = 1))
  
    if(!legend) {
   p <- p + theme(legend.position = "none")
  }
    print(p)}
  
plot_O_pub_Str2 <- function(prediction, m_code, legend = TRUE) {
  p<- ggplot() +
  geom_tile(
    data = prediction,
    aes(x = Lon, y = Lat, fill = omega_s),
    width = tile_width,  # adjust to your grid spacing
    height = tile_height ) +
  scale_fill_gradient2(name = expression(omega[s]),
     low = "red", mid = "white", high = "blue",
    midpoint = 0,
    limits = color_limits_Str2) + 
    labs(x = "Longitude", y = "Latitude", title = m_code) +
    theme(panel.background = element_rect(fill = "gray85", color = NA))+
    geom_sf(data = coast, fill = "gray20", col = "gray50") +
    coord_sf(xlim = c(22.5, 25), ylim = c(38.3, 40.9))+
    scale_x_continuous(breaks = seq(22.5, 25, by = 1))
  
if(!legend) {
   p <- p + theme(legend.position = "none")
  }
    print(p)
}


#Visualize the spatial fields
plotomega_Str1_06 <- plot_O_pub_Str1(model_M3Str1_06_pred2$data, "2006", legend = TRUE) 
plotomega_Str2_06 <- plot_O_pub_Str2(model_M3Str2_06_pred2$data, "2006", legend = TRUE) 
```

## Model 4: GLMM with Tweedie and spatial autocorralation

### Write M4 function 
```{r}
fit_glm_TWsp_M4 <- function(data,PA_polygone,bspde,plot_basic) {
 
# Get data frame name
df_name <- deparse(substitute(data))
model_name <- paste0("model_M4", df_name)
Survey <- as.character(data$Survey[1])

data <- data %>% filter(!is.na(Abundance))
data <- data %>% filter(!is.na(w_sn))
Year <- as.character(data$Year[1])

# Build dynamic object names
  plot_name  <- paste0("plot_M4_", df_name)
  plot_log_name  <- paste0("plot_log_M4_", df_name)
  omega_name  <- paste0("omegaSD_M4_", df_name)
  range_name  <- paste0("range_M4_", df_name)
  dispersion_name  <- paste0("dispersion_M4_", df_name)

  P_est_name <- paste0("P_est_M4_", df_name)
  z_est_name <- paste0("z_est_M4_", df_name)
  
  P_se_name  <- paste0("P_se_M4_", df_name)
  z_se_name  <- paste0("z_se_M4_", df_name)
  
  P_ci_name  <- paste0("P_ci_M4_", df_name)
  z_ci_name  <- paste0("z_ci_M4_", df_name)
  
# Fit model
model4 <-  sdmTMB(
  Abundance ~ 1 + (Age),
  data = data,
  family = tweedie(link = "log"),
  mesh = bspde,
  spatial = "on")  
assign(model_name, model4, envir = .GlobalEnv)

#Summary
summary(model4)
sanity(model4)

params <-tidy(model4, conf.int = TRUE)
tidy(model4, effects = "ran_par", conf.int = TRUE)

# Extract param
 P_est <- exp(params$estimate[1])
 z_est <- -(params$estimate[2]) 
  
# SEs
  P_se_log <-  params$std.error[1]
  z_se <- params$std.error[2]
  
# CIs 
  P_ci <- NA
  z_ci <-  c(-params$conf.high[2], -params$conf.low[2])  
  
# Compute transformed estimates
#LogP0 
CV_P <- sqrt(exp(P_se_log^2) - 1)
P_se <- P_est * CV_P
P_ci_lower <- exp(params$conf.low[1]) 
P_ci_upper <- exp(params$conf.high[1])

# Spatial stracture
  rep <- model4$tmb_obj$report()
  sigma_O <- rep$sigma_O [1,1]
  range<-  rep$range[1,1]
 dispersion <-model4$sd_report$value["phi"]%>% as.numeric()

  assign(omega_name, sigma_O, envir = .GlobalEnv)
  assign(range_name, range, envir = .GlobalEnv)
  assign(dispersion_name, dispersion, envir = .GlobalEnv)

outablePo <- data.frame(
    dataset       =  Survey,
    parameter     = c("P"),
    estimate_log  = params$estimate[1], 
    estimate      = c(P_est),       # Unlog
    se            = c(P_se) ,       #se of the estimated model,
    ci_lower      = c(P_ci_lower),
    ci_upper      = c(P_ci_upper),
    CV            = c(CV_P),        
    model         = "M4",
    stringsAsFactors = FALSE) 

  outableZ <- data.frame(
    dataset   = c(Survey),
    parameter = "z",
    estimate  = c(z_est),
    se        = c(z_se),
    ci_lower  = c(z_ci[1]), #It haves been reversed because z insted of -z is modeld
    ci_upper  = c(z_ci[2]),
    CV        = c(z_se/z_est),
    model     = "M4",
    stringsAsFactors = FALSE)
  
#Predict in data points 
# Back predict on the data:  for the  omega -and the  the plots 
predic.grid1 <- data %>% filter(!is.na(Stage))  %>% filter(!is.na(Age))  
predict1 <- predict(model4,newdata=predic.grid1,re_form = NA) #Turn of the spatial filed in the prediction

model_prediction1 <- paste0("model_M4", df_name, "_pred1")
assign(model_prediction1, predict1, envir = .GlobalEnv)

# Plot
  plot <- 
     plot_basic +
   geom_line(data = predict1, 
             aes(x=Age, y= exp(est)), 
             color = brewer.pal(7, "Dark2")[5], linewidth = 1) +
   labs(title = "",
        subtitle = paste(df_name),
        x = "Age",
        y = "Abundance"  ) +
    theme_bw() +
    theme(panel.grid.major = element_line(color = "white"))
  
assign(plot_name, plot, envir = .GlobalEnv) 

#For the Ptot Prediction - as index standardization
#1. Create the extrapolation grid (PA)
grid_1km <- st_make_grid(
  PA_polygone,
  cellsize = 1,        # 1 unit = 1 km
  square = TRUE,
  what = "polygons")

grid_clipped <- st_intersection(grid_1km, PA_polygone)  # Do the grid only in the Positive area 

plot(grid_clipped)
centroids <- st_centroid(grid_clipped)
centroids_df <- st_coordinates(centroids) %>%
  as.data.frame() 

centroids_m <- centroids %>% st_sf()  # ensure sf
st_geometry(centroids_m) <- st_geometry(centroids_m) * 1000  # km -> m
st_crs(centroids_m) <- 32635
centroids_wgs84 <- st_transform(centroids_m, crs = 4326)

lonlat_df <- st_coordinates(centroids_wgs84) %>%
  as.data.frame() %>% dplyr::rename(Lon = X, Lat = Y)

centroids_df <- bind_cols(centroids_df, lonlat_df)

#This is the new prediction grid that will force the prediction index standardization 
#log(Pij)=log(Po) + ω +  (z*t) , but z*t=0 so --> log(Pij)=log(Po) + ω (Prediction is now Ptot)
predic.grid2<- centroids_df %>% mutate(Age=0)

#Predict
predict2 <- predict(model4,newdata=predic.grid2,return_tmb_object = TRUE) 
model_prediction2 <- paste0("model_M4", df_name, "_pred2")
assign(model_prediction2, predict2, envir = .GlobalEnv)
 
#3.Get the index (Ptot)
index <- get_index(predict2, area =10^6, bias_correct = TRUE)  # 10^6 for making kM1--> M1
index <- index %>% mutate(cv = sqrt(exp(se^2) - 1))   #Since se refere ot the log estimate

outablelPtot <- data.frame(
  dataset       =  Survey,
  parameter = "Ptot",
  estimate  = index$est, #estimate in the original scale
  se        =  index$cv*index$est,  # re-estimate it because it was in log(abundance)
  ci_lower  =  index$lwr,
  ci_upper  =  index$upr,
  CV        = index$cv,         # re-estimate it because it was in log(abundance))
   model         = "M4",
   stringsAsFactors = FALSE) 

# Output tables
return(list(outablePo= outablePo, outableZ=outableZ,outablelPtot=outablelPtot))
}
```

### Fit M4
```{r}
results_all_M4 <- bind_rows(
fit_glm_TWsp_M4(Str1_06,PA_2006_Str1,bspdeStr1_06,plot_M3_Str1_06),
fit_glm_TWsp_M4(Str2_06,PA_2006_Str2,bspdeStr2_06,plot_M3_Str2_06))
```

### Get P,z results and estimate Ptot
```{r}
outablePo_temp_Str1<- results_all_M4$outablePo %>% filter(grepl("^Str1_",dataset)) %>% mutate(across(where(is.numeric), ~ round(.x, 3)))
outablePo_temp_Str1 <- outablePo_temp_Str1 %>% select (-estimate_log)
Str1_resultsPo <- rbind(Str1_resultsPo,outablePo_temp_Str1)
gt(Str1_resultsPo)

outablePo_temp_Str2<- results_all_M4$outablePo %>% filter(grepl("^Str2",dataset)) %>% mutate(across(where(is.numeric), ~ round(.x, 3)))
outablePo_temp_Str2 <- outablePo_temp_Str2 %>% select (-estimate_log)
Str2_resultsPo <- rbind(Str2_resultsPo,outablePo_temp_Str2)
gt(Str2_resultsPo)

outableZ_temp_Str1 <- results_all_M4$outableZ %>% filter(grepl("^Str1_",dataset)) %>% mutate(across(where(is.numeric), ~ round(.x, 3)))
Str1_resultsZ <- rbind(Str1_resultsZ,outableZ_temp_Str1)
gt(Str1_resultsZ)

outableZ_temp_Str2 <- results_all_M4$outableZ %>% filter(grepl("^Str2_",dataset)) %>% mutate(across(where(is.numeric), ~ round(.x, 3)))
Str2_resultsZ <- rbind(Str2_resultsZ,outableZ_temp_Str2)
gt(Str2_resultsZ)


Str1_resultsPtot_temp <-results_all_M4$outablelPtot %>% filter(grepl("^Str1_",dataset)) 

Str1_resultsPtot <- rbind(Str1_resultsPtot,Str1_resultsPtot_temp)
gt(Str1_resultsPtot)

Str2_resultsPtot_temp <-results_all_M4$outablelPtot %>% filter(grepl("^Str2_",dataset)) 

Str2_resultsPtot <- rbind(Str2_resultsPtot,Str2_resultsPtot_temp)
gt(Str2_resultsPtot)

```


### Summary plots
```{r}
# Eastern stratum Po plot
plot_Po_Str1<-
ggplot() +
  geom_point(data = Str1_resultsPo, aes(y = estimate, x = dataset, col = model),
             size = 2,position = position_dodge(width = 0.4)) + 
  geom_errorbar(data = Str1_resultsPo, 
                aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model), 
                width = 0.2,linewidth = 1, position = position_dodge(width = 0.4)) +  
  theme_bw() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[0] ~ (eggs ~ m^-2 ~ day^-1))) +xlab("")+
   scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))
  
# Western stratum Po plot
plot_Po_Str2<-
ggplot() +
  geom_point(data = Str2_resultsPo, 
             aes(y = estimate, x = dataset, col = model),
             size = 2, position= position_dodge(width = 0.4))  + 
  geom_errorbar(data = Str2_resultsPo, 
                aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model), 
                width = 0.2,linewidth = 1,position=position_dodge(width = 0.4))  +  
  theme_bw() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[0] ~ (eggs ~ m^-2 ~ day^-1))) +xlab("")+
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))
  
# Eastern stratum Z plot
plot_z_Str1<-
ggplot() +
  geom_point(data = Str1_resultsZ,
             aes(y = estimate, x = dataset, col = model),
             size = 2, position=position_dodge(width = 0.4)) + 
  geom_errorbar(
    data = Str1_resultsZ, 
    aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model), 
    width = 0.2,linewidth = 1, position=position_dodge(width = 0.4)) +  
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression (Z~(day^-1))) +xlab("")+
   scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))
  
# Western stratum Z plot
plot_z_Str2<-
ggplot() +
  geom_point(
    data = Str2_resultsZ, 
    aes(y = estimate, x = dataset, col = model),
    size =2, position=position_dodge(width = 0.4)) + 
 geom_errorbar(data = Str2_resultsZ,
               aes(x = dataset, ymin = ci_lower, ymax = ci_upper, color = model), 
               width = 0.2,linewidth = 1,position=position_dodge(width = 0.4))  +  
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression (Z~(day^-1))) +xlab("")+
   scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))
  
# Eastern stratum Ptot plot
plot_Ptot_Str1<-
ggplot() +
  geom_point(data = Str1_resultsPtot, 
             aes(y = estimate/10^11, x = dataset, col = model),
             size=2,position=position_dodge(width = 0.4))+ 
  geom_errorbar(data = Str1_resultsPtot, 
                aes(x= dataset, ymin = ci_lower /10^11, ymax = ci_upper/10^11, color = model),
                width = 0.2,linewidth = 1, position=position_dodge(width = 0.4))+   
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[tot] ~ (eggs ~ day^-1) ~x10^11 )) +xlab("")+
  scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))
  
# Western stratum Ptot plot
plot_Ptot_Str2<-
ggplot() +
  geom_point(data = Str2_resultsPtot,
             aes(y = estimate/10^11, x = dataset, col = model),
             size=2, position=position_dodge(width = 0.4)) + 
  geom_errorbar(data = Str2_resultsPtot, 
                aes(x= dataset, ymin = ci_lower /10^11, ymax = ci_upper/10^11, color = model), 
                width = 0.2,linewidth = 1, position=position_dodge(width = 0.4)) +   
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[tot] ~ (eggs ~ day^-1) ~x10^11 )) +xlab("")+
    scale_fill_manual(name = "Model",brewer.pal(1, "Set1"))
  
Summary_Plots<- 
  print(plot_grid(
    plot_Po_Str1,plot_Po_Str2,
    plot_z_Str1, plot_z_Str2, 
    plot_Ptot_Str1, plot_Ptot_Str2,
    ncol =  2))
```

### Fits
```{r}
Ploted_fits_Str1<- print(plot_grid(plot_M4_Str1_06))
Ploted_fits_Str2<- print(plot_grid(plot_M4_Str2_06))
```

### QQ and Residuals 
```{r}
Residuals_plots<- function(data,model,model_name){
    df_name <- deparse(substitute(data))
    res_plot_name  <- paste0("res_plot_",model_name,df_name)
    res_qq_name  <- paste0("res_qq_",model_name,df_name)
    data <- data %>% filter(!is.na(Abundance), !is.na(w_sn))
     year <- sub(".*_(\\d{2}).*", "\\1", df_name) ;year <- paste0("20", year)
   
# Determine stratum
 stratum <- case_when(
  grepl("^Str1", df_name) ~ "Eastern stratum",
  grepl("^Str2", df_name) ~ "Western stratum",
  TRUE ~ "")

# Generate predictions 
df_res <- data.frame(
  Age = data$Age,
  Predicted = exp((predict(model, newdata = data.frame(Age = data$Age,X=data$X,Y=data$Y)))$est), # Unlog
  Observed= data$Abundance) %>% mutate (Residual= residuals(model,type ="mle-mvn"))

#Pearson Cor  
r_val <- cor(df_res$Predicted, df_res$Observed, use = "complete.obs",method = "pearson")

# Residual vs Fitted plot
p_obs_pred <- ggplot(df_res, aes(x = Predicted, y = Observed)) +
  geom_point(alpha = 0.6, color=brewer.pal(7, "Dark2")[5]) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  theme_bw() +
  labs(
    x = "",
    y = "Observed",
    subtitle = paste(model_name) )+
annotate(
  "text",
  x = -Inf,
  y = Inf,
  label = sprintf("r = %.3f", r_val),
  hjust = -0.1,
  vjust = 1.1,
  size =4, color="red") + 
  theme_bw() +
  theme(plot.margin = margin(t=0,r = 1,b =0,l = 1))
  
p_res_fit <- ggplot(df_res, aes(x = Predicted, y = Residual)) +
    geom_point(alpha = 0.6,color=brewer.pal(7, "Dark2")[5]) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    theme_bw() +
    labs(
      x = "",
      y = "Residuals",
      subtitle="")+
    theme_bw() +
    theme(plot.margin = margin(t = 0,r = 1,b =0,l = 1)) 

p_qq <- ggplot(df_res, aes(sample = Residual)) +
    stat_qq(color=brewer.pal(7, "Dark2")[5]) +
    stat_qq_line(linetype = "dashed", color = "red") +
    theme_bw() +
    labs(
      x = "",
      y = "Sample Q",
      subtitle="") 
    theme_bw() +
    theme(plot.margin = margin(t=0,r = 1,b =0,l = 1)) 

res_plots<- plot_grid(p_obs_pred,p_res_fit,p_qq, ncol=3)
assign(res_plot_name, res_plots, envir = .GlobalEnv) 
assign(res_qq_name, res_plots, envir = .GlobalEnv)  

}
Residuals_plots(Str1_06,model_M4Str1_06, "M4") ; res_plot_M4Str1_06 ;res_qq_M4Str1_06
Residuals_plots(Str2_06,model_M4Str2_06, "M4") ; res_plot_M4Str2_06 ;res_qq_M4Str2_06

Residuals_Str1_06<- print(plot_grid(res_plot_M0Str1_06,res_plot_M1Str1_06,res_plot_M2Str1_06,res_plot_M3Str1_06,res_plot_M4Str1_06,nrow=5))
Residuals_Str2_06<- print(plot_grid(res_plot_M0Str2_06,res_plot_M1Str2_06,res_plot_M2Str2_06,res_plot_M3Str2_06,res_plot_M3Str2_06,nrow=5))
```
### DHAMR residuals 
```{r}
set.seed(123)
Residuals_test <- data.frame(
    Model = character(),
    Uniformity = character(),
    Uniformity_p = numeric(),
    Dispersion = character(),
    Dispersion_p= numeric(),
    Quantile   = character(),
    Quantile_p = numeric(),
    ZeroInfl   = character(),
    ZeroInfl_p = numeric(),
    SpatialAut = character(),
    SpatialAut_p= numeric())

resid_Chek <- function(model, data, a = 0.05, model_name = NULL) {
  
  df_name <- deparse(substitute(data))
  model_name <- paste0("model_M4", df_name)
  
 # Clean data
 data <- data %>% filter(!is.na(Abundance), !is.na(w_sn))

 # Simulate and calculate residuals
 s_fit <- simulate(model, nsim = 500, type = "mle-mvn")
 r_fit <- dharma_residuals(s_fit, model, return_DHARMa = TRUE)
  
 # Run DHARMa tests

 res_tests<-DHARMa::testResiduals(r_fit,plot=T)  
 q_test   <- DHARMa::testQuantiles(r_fit) # same as plotResiduals quantile residuals versus (ranked) fitted values. Check for homogeneity of variance-
 zi_test  <- DHARMa::testZeroInflation(r_fit)

# For the spatial autocoralation 
simulationOutputS <- recalculateResiduals(
  r_fit,
  group = paste(data$X, data$Y),
  aggregateBy = mean)
  
sp_test <- DHARMa::testSpatialAutocorrelation(
  simulationOutputS,
  x = unique(data$X),
  y = unique(data$Y))
  
# Decide pass/fail
un_re  <- ifelse(res_tests$uniformity$p.value  < a, "Fail", "Pass")
dis_re  <- ifelse(res_tests$dispersion$p.value  < a, "Fail", "Pass")
q_res  <- ifelse(q_test$p.value  < a, "Fail", "Pass")
zi_res <- ifelse(zi_test$p.value < a, "Fail", "Pass")
sp_res <- ifelse(sp_test$p.value < a, "Fail", "Pass")
  
# Add results to global table
new_row <- data.frame(
  Model = paste(model_name),
  Uniformity = un_re,
  Uniformity_p = round(res_tests$uniformity$p.value,4),
  Dispersion = dis_re,
  Dispersion_p = round(res_tests$dispersion$p.value,4),
  Quantile = q_res,
  Quantile_p= round(q_test$p.value,4),
  ZeroInfl = zi_res,
  ZeroInfl_p= round(zi_test$p.value,4),
  SpatialAuto = sp_res,
  SpatialAuto_p= round(sp_test$p.value,4) )
  
assign("Residuals_test", rbind(get("Residuals_test", envir = .GlobalEnv), new_row), envir = .GlobalEnv)

}

resid_Chek(model_M4Str1_06,Str1_06)
resid_Chek(model_M4Str2_06,Str2_06)

gt(Residuals_test,caption = "Residuals Test")
Residuals_test_M4 <- Residuals_test
# Do a summary 
# Select only the test columns
test_cols <- c("Uniformity", "Dispersion", "Quantile", "ZeroInfl", "SpatialAuto")

# Function to count Pass
count_pass <- function(x) sum(x == "Pass")

# Create summary
summary_table <- data.frame(
  Test = test_cols,
  Pass = sapply(Residuals_test[test_cols], count_pass),
  Total = nrow(Residuals_test))

# Add Pass Rate as percentage
summary_table$PassRate <- round(summary_table$Pass / summary_table$Total * 100, 1)

# View the summary table
print(summary_table)
Residuals_test_M4_sum <- summary_table
```

### Omegas
```{r}
#plot all the omegas 
all_omega_Str1 <- c(
  model_M4Str1_06_pred2$data$omega_s)

color_limits_Str1 <- c(min(all_omega_Str1, na.rm = TRUE),max(all_omega_Str1, na.rm = TRUE)) 

all_omega_Str2 <- c( model_M4Str2_06_pred2$data$omega_s)

color_limits_Str2 <- c(min(all_omega_Str2, na.rm = TRUE),max(all_omega_Str2, na.rm = TRUE)) 

#I do this with tiles as my surface is continuous.
tile_size <- 0.015
tile_width <- 1 / (111 * cos(25 * pi/180))  # ~0.0102°
tile_height <- 1 / 111  # ~0.0090°
tile_height <- tile_height * 1.15
tile_width <- tile_width * 1.2

plot_O_pub_Str1 <- function(prediction, m_code, legend = TRUE) {
  
p<- ggplot() +
   # Coastline
  geom_tile(
    data = prediction,
    aes(x = Lon, y = Lat, fill = omega_s),
    width = tile_width,  # adjust to your grid spacing
    height = tile_height
  ) +
   scale_fill_gradient2(name = expression(omega[s]),
     low = "red", mid = "white", high = "blue",
     midpoint = 0,
     limits = color_limits_Str1)+
    labs(x = "Longitude", y = "Latitude", title = m_code) +
    theme(panel.background = element_rect(fill = "gray85", color = NA))+
    geom_sf(data = coast, fill = "gray20", col = "gray50") +
    coord_sf(xlim = c(23.5, 26), ylim = c(39.3, 41)) +
    scale_x_continuous(breaks = seq(23.5, 26.5, by = 1))
     
 if(!legend) {
   p <- p + theme(legend.position = "none")
  }
    print(p)
}


plot_O_pub_Str2 <- function(prediction, m_code, legend = TRUE) {
  
p<- ggplot() +
  geom_tile(
    data = prediction,
    aes(x = Lon, y = Lat, fill = omega_s),
    width = tile_width,  
    height = tile_height
  ) +
    scale_fill_gradient2(name = expression(omega[s]),
     low = "red", mid = "white", high = "blue",
     midpoint = 0,
     limits = color_limits_Str2)+
  labs(x = "Longitude", y = "Latitude", title = m_code) +
    theme(panel.background = element_rect(fill = "gray85", color = NA))+
    geom_sf(data = coast, fill = "gray20", col = "gray50") +
    coord_sf(xlim = c(22.5, 25), ylim = c(38.3, 40.9))+
    scale_x_continuous(breaks = seq(22.5, 25, by = 1))

if(!legend) {
   p <- p + theme(legend.position = "none")
  }
    print(p)
}

plotomega_Str1_06 <- plot_O_pub_Str1(model_M4Str1_06_pred2$data, "2006", legend = TRUE) 
plots_grid_Str1   <- plot_grid(plotomega_Str1_06)

plotomega_Str2_06 <- plot_O_pub_Str2(model_M4Str2_06_pred2$data, "2006", legend = TRUE) 
plots_grid_Str2   <- plot_grid(plotomega_Str2_06)
```


## Model 5: GLMM with tweedie, spatial autocorralation and spatial z

### Write M5 function 

```{r}
fit_glm_TW_svc_M5 <- function(data,PA_polygone,bspde,plot_basic)  {

# Get data frame name
df_name <- deparse(substitute(data))
model_name <- paste0("model_M5", df_name)
Survey <- as.character(data$Survey[1])
  
data <- data %>% filter(!is.na(Abundance))
data <- data %>% filter(!is.na(w_sn))
Year <- as.character(data$Year[1])
  
# Build dynamic object names
plot_name  <- paste0("plot_M5_", df_name)
plot_log_name  <- paste0("plot_log_M5_", df_name)
omega_name  <- paste0("omegaSD_M5_", df_name)
z_name  <- paste0("svcSD_M5_", df_name)
range_name  <- paste0("range_M5_", df_name)
dispersion_name  <- paste0("dispersion_M5_", df_name)

P_est_name <- paste0("P_est_M5_", df_name)
z_est_name <- paste0("z_est_M5_", df_name)
  
P_se_name  <- paste0("P_se_M5_", df_name)
z_se_name  <- paste0("z_se_M5_", df_name)
  
P_ci_name  <- paste0("P_ci_M5_", df_name)
z_ci_name  <- paste0("z_ci_M5_", df_name)

model5 <-  sdmTMB(
  Abundance ~ 1 + (Age),
  data = data,
  family = tweedie (link = "log"),
  #weights = data$w_sn,
  mesh = bspde,
  spatial = "on",
  spatial_varying = ~ Age)

assign(model_name, model5, envir = .GlobalEnv)

#Summary
summary(model5)
sanity(model5)

params <-tidy(model5, conf.int = TRUE)
tidy(model5, effects = "ran_par", conf.int = TRUE)

 # Extract params
  P_est <- exp(params$estimate[1]) 
  z_est <- -(params$estimate[2])
  
  # SEs
  P_se_log <-  params$std.error[1]
  z_se <- params$std.error[2]
  
  # CIs 
  P_ci <- NA
  z_ci <-  c(-params$conf.high[2], -params$conf.low[2])
  
  # Compute transformed estimates
  #LogP0  
  CV_P <- sqrt(exp(P_se_log^2) - 1)
  P_se <- P_est * CV_P
  P_ci_lower <- exp(params$conf.low[1]) 
  P_ci_upper <- exp(params$conf.high[1])

  # Spatial stracture
  rep <- model5$tmb_obj$report()
  sigma_O <- rep$sigma_O [1,1]
  sigma_Z <- rep$sigma_Z [1,1]
  range  <-  rep$range[1,1]
  dispersion <-model5$sd_report$value["phi"]%>% as.numeric()

  assign(omega_name, sigma_O, envir = .GlobalEnv)
  assign(range_name, range, envir = .GlobalEnv)
  assign(z_name, sigma_Z, envir = .GlobalEnv)
  assign(dispersion_name, dispersion, envir = .GlobalEnv)

  outablePo <- data.frame(
    dataset       =  Survey,
    parameter     = c("P"),
    estimate_log  = params$estimate[1], 
    estimate      = c(P_est),       # Unlog
    se            = c(P_se) ,       #se of the estimated model,
    ci_lower      = c(P_ci_lower),
    ci_upper      = c(P_ci_upper),
    CV            = c(CV_P),        
    model         = "M5",
    stringsAsFactors = FALSE) 
  
  outableZ <- data.frame(
    dataset   = c(Survey),
    parameter = "z",
    estimate  = c(z_est),
    se        = c(z_se),
    ci_lower  = c(z_ci[1]), #It haves been reversed because z insted of -z is modeld
    ci_upper  = c(z_ci[2]),
    CV        = c(z_se/z_est),
    model     = "M5",
    stringsAsFactors = FALSE)

 #Predict in data points 
  # Back predict on the data:  for the  omega -and the  the plots 
predic.grid1 <- data %>% filter(!is.na(Stage))  %>% filter(!is.na(Age))  
predict1 <- predict(model5,newdata=predic.grid1,re_form = NA) #Turn of the spatial filed in the prediction
  
model_prediction1 <- paste0("model_M5", df_name, "_pred1")
assign(model_prediction1, predict1, envir = .GlobalEnv)
  
# Plot
plot <- 
    plot_basic +
     geom_line(data = predict1, aes(x=Age, y= exp(est)), color = brewer.pal(7, "Dark2")[6], linewidth = 1) +
    labs(
      title = "",
      subtitle = paste(df_name),#, ": P=", round(P_est, 3), ", z=", round(z_est, 3)),
      x = "Age",
      y = "Abundance"
    ) +
    theme_bw() +
    theme(panel.grid.major = element_line(color = "white"))
  
  assign(plot_name, plot, envir = .GlobalEnv) #print(plo)
  
 
  #For the Ptot Prediction -Treet as index stundardization
  #1. Create the extrapolation grid (PA- division)
  grid_1km <- st_make_grid(
    PA_polygone,
    cellsize = 1,        # 1 unit = 1 km
    square = TRUE,
    what = "polygons")
  
  grid_clipped <- st_intersection(grid_1km, PA_polygone)  # Do the grid only in the Positive area 
  
  plot(grid_clipped)
  centroids <- st_centroid(grid_clipped)
  centroids_df <- st_coordinates(centroids) %>%
    as.data.frame() 
  
  centroids_m <- centroids %>% st_sf()  # ensure sf
  st_geometry(centroids_m) <- st_geometry(centroids_m) * 1000  # km -> m
  st_crs(centroids_m) <- 32635
  centroids_wgs84 <- st_transform(centroids_m, crs = 4326)
  
  lonlat_df <- st_coordinates(centroids_wgs84) %>%
    as.data.frame() %>% dplyr::rename(Lon = X, Lat = Y)
  
  centroids_df <- bind_cols(centroids_df, lonlat_df)
  
  #This is the new prediction grid that will force the prediction index standardization 
  #log(Pij)=log(Po) + ω +  (z*t) , but z*t=0 so --> log(Pij)=log(Po) + ω (Prediction is now Ptot)
  predic.grid2<- centroids_df %>% mutate(Age=0)
  
  #Predict
  predict2 <- predict(model5,newdata=predic.grid2,return_tmb_object = TRUE) 
  model_prediction2 <- paste0("model_M5", df_name, "_pred2")
  assign(model_prediction2, predict2, envir = .GlobalEnv)
  
  #3.Get the index (Ptot)
  index <- get_index(predict2, area =10^6, bias_correct = TRUE)  # 10^6 for making kM1--> M1
  index <- index %>% mutate(cv = sqrt(exp(se^2) - 1))   #Since se refere ot the log estimate
  
  outablelPtot <- data.frame(
    dataset       =  Survey,
    parameter = "Ptot",
    estimate  = index$est, #estimate in the original scale
    se        =  index$cv*index$est,  # re-estimate it because it was in log(abundance)
    ci_lower  =  index$lwr,
    ci_upper  =  index$upr,
    CV        = index$cv,          # re-estimate it because it was in log(abundance))
    model         = "M5",
    stringsAsFactors = FALSE) 
  
  # Output tables
  return(list(outablePo= outablePo, outableZ=outableZ,outablelPtot=outablelPtot))
}
```
### Fit M5
```{r}
results_all_M5 <- bind_rows(
fit_glm_TW_svc_M5(Str1_06,PA_2006_Str1,bspdeStr1_06,plot_M4_Str1_06),
fit_glm_TW_svc_M5(Str2_06,PA_2006_Str2,bspdeStr2_06,plot_M4_Str2_06))
# (Z=0)

```


### Get P,z results and estimate Ptot
```{r}
all_datasets_Str1 <- data.frame(
  dataset = c( "Str1_03","Str1_04","Str1_05","Str1_06","Str1_08","Str1_14","Str1_16","Str1_19","Str1_22"))

all_datasets_Str2 <- data.frame(
  dataset = c( "Str2_03","Str2_04","Str2_05","Str2_06","Str2_08","Str2_14","Str2_16","Str2_19","Str2_22"))

outablePo_temp_Str1 <- results_all_M5$outablePo %>% 
  filter(grepl("^Str1_", dataset)) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  select(-estimate_log) %>%
  right_join(all_datasets_Str1, by = "dataset") %>%
  arrange(dataset) %>%
  mutate(model = "M5")
Str1_resultsPo <- rbind(Str1_resultsPo,outablePo_temp_Str1)
gt(Str1_resultsPo)

outablePo_temp_Str2 <- results_all_M5$outablePo %>%
  filter(grepl("^Str2_", dataset)) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  select(-estimate_log) %>%
  right_join(all_datasets_Str2, by = "dataset") %>%
  arrange(dataset) %>%
  mutate(model = "M5")
Str2_resultsPo <- rbind(Str2_resultsPo,outablePo_temp_Str2)
gt(Str2_resultsPo)

outableZ_temp_Str1 <- results_all_M5$outableZ   %>% 
  filter(grepl("^Str1_",dataset)) %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  right_join(all_datasets_Str1, by = "dataset") %>%
  arrange(dataset) %>%
  mutate(model = "M5")

Str1_resultsZ <- rbind(Str1_resultsZ,outableZ_temp_Str1)
gt(Str1_resultsZ)

outableZ_temp_Str2 <- results_all_M5$outableZ %>%
  filter(grepl("^Str2_",dataset)) %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
 right_join(all_datasets_Str2, by = "dataset") %>%
  arrange(dataset) %>%
  mutate(model = "M5")
 Str2_resultsZ <- rbind(Str2_resultsZ,outableZ_temp_Str2)
gt(Str2_resultsZ)

#Plot will be prediction output 
Str1_resultsPtot_temp <-results_all_M5$outablelPtot  %>% filter(grepl("^Str1_",dataset)) %>%  
 mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
 right_join(all_datasets_Str1, by = "dataset") %>%
 arrange(dataset) %>%
 mutate(model = "M5")

Str1_resultsPtot <- rbind(Str1_resultsPtot,Str1_resultsPtot_temp)
gt(Str1_resultsPtot)

Str2_resultsPtot_temp <-results_all_M5$outablelPtot %>% filter(grepl("^Str2_",dataset)) %>%  
 mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
 right_join(all_datasets_Str2, by = "dataset") %>%
 arrange(dataset) %>%
 mutate(model = "M5")

Str2_resultsPtot <- rbind(Str2_resultsPtot,Str2_resultsPtot_temp)
gt(Str2_resultsPtot)
```

### Fits 
```{r}
Ploted_fits_Str1<- print(plot_grid(plot_M5_Str1_06))
```

### QQ and Residuals
```{r}
Residuals_plots<- function(data,model,model_name){
    df_name <- deparse(substitute(data))
    res_plot_name  <- paste0("res_plot_",model_name,df_name)
    res_qq_name  <- paste0("res_qq_",model_name,df_name)
    data <- data %>% filter(!is.na(Abundance), !is.na(w_sn))
    year <- sub(".*_(\\d{2}).*", "\\1", df_name) ;year <- paste0("20", year)
   
 # Determine stratum
 stratum <- case_when(
  grepl("^Str1", df_name) ~ "Eastern stratum",
  grepl("^Str2", df_name) ~ "Western stratum",
  TRUE ~ "")
 
# Generate predictions on the original Age values
df_res <- data.frame(
  Age = data$Age,
  Predicted = exp((predict(model, newdata = data.frame(Age = data$Age,X=data$X,Y=data$Y)))$est),# Unlog
  Observed= data$Abundance) %>% mutate (Residual= residuals(model,type ="mle-mvn"))

#Pearson Cor  
r_val <- cor(df_res$Predicted, df_res$Observed, use = "complete.obs",method = "pearson")

# Residual vs Fitted plot
p_obs_pred <- ggplot(df_res, aes(x = Predicted, y = Observed)) +
  geom_point(alpha = 0.6, color = brewer.pal(7, "Dark2")[6]) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(
    x = "Fitted",
    y = "Observed",
    subtitle = paste(model_name) )+
annotate(
  "text",
  x = -Inf,
  y = Inf,
  label = sprintf("r = %.3f", r_val),
  hjust = -0.1,
  vjust = 1.1,
  size =4, color="red")  +
  theme_bw() +
  theme(plot.margin = margin(t=0.2,r = 1,b =0,l = 1))
  
p_res_fit <- ggplot(df_res, aes(x = Predicted, y = Residual)) +
    geom_point(alpha = 0.6, color = brewer.pal(7, "Dark2")[6]) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    theme_bw() +
    labs(
      x = "Fitted",
      y = "Residuals", 
      subtitle="") + 
      theme_bw() +
  theme(plot.margin = margin(t=0.2,r = 1,b =0,l = 1))

p_qq <- ggplot(df_res, aes(sample = Residual)) +
    stat_qq( color = brewer.pal(7, "Dark2")[6]) +
    stat_qq_line(linetype = "dashed", color = "red") +
    theme_bw() +
    labs(
      x = "Theoretical Q",
      y = "Sample Q",
      subtitle="")  + 
      theme_bw() +
     theme(plot.margin = margin(t=0.2,r = 1,b =0,l = 1))

res_plots<- plot_grid(p_obs_pred,p_res_fit,p_qq, ncol=3)
assign(res_plot_name, res_plots, envir = .GlobalEnv) 
assign(res_qq_name, res_plots, envir = .GlobalEnv)  
}


Residuals_plots(Str1_06,model_M5Str1_06, "M5") ; res_plot_M5Str1_06 ;res_qq_M5Str1_06
#Residuals_plots(Str2_06,model_M5Str2_06, "M5") ; res_plot_M5Str2_06 ;res_qq_M5Str2_06

Residuals_Str1_06 <- (  
  res_plot_M0Str1_06 /
  res_plot_M1Str1_06 /
  res_plot_M2Str1_06 /
  res_plot_M3Str1_06 /
  res_plot_M4Str1_06 /
  res_plot_M5Str1_06) + plot_annotation(title = "Eastern Stratum 2006")

Residuals_Str2_06 <- (  
  res_plot_M0Str2_06 /
  res_plot_M1Str2_06 /
  res_plot_M2Str2_06 /
  res_plot_M3Str2_06 /
  res_plot_M4Str2_06 /
  plot_spacer()) + plot_annotation(title = "Western Stratum 2006")

```

### Summary plots
```{r}
Str1_resultsPo$dataset_short <- paste0("20", sub("^Str1_", "", Str1_resultsPo$dataset))
  
plot_Po_Str1<-
ggplot() +
  geom_point(data = Str1_resultsPo,
             aes(y = estimate, x = dataset_short, col = model),
             size = 1,position = position_dodge(width = 0.65)) + 
  geom_errorbar(data = Str1_resultsPo, 
                aes(x = dataset_short, ymin = ci_lower, ymax = ci_upper, color = model), 
                width = 0.2,linewidth = 0.5, position = position_dodge(width = 0.65)) +  
  theme_bw() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("")+ ylab("")+
    scale_colour_brewer(name = "Model", palette = "Dark2")+theme(legend.position = "none")    +   facet_wrap(~factor("B. Eastern Stratum"))+
  theme(strip.background = 
         element_rect(fill = "white",colour = "black",strip.background = element_blank(),strip.text = element_text(face = "bold"))) 
  

Str1_resultsPo <- Str1_resultsPo %>% select(-dataset_short)

Str2_resultsPo$dataset_short <- paste0("20", sub("^Str2_", "", Str2_resultsPo$dataset))

plot_Po_Str2<-ggplot() +
geom_point(data = Str2_resultsPo, 
           aes(y = estimate, x = dataset_short, col = model),
           size = 1, position=    position_dodge(width = 0.65))  + 
  geom_errorbar(data = Str2_resultsPo,
                aes(x = dataset_short, ymin = ci_lower, ymax = ci_upper, color = model), 
                width = 0.2,linewidth = 0.5,position=position_dodge(width = 0.65))  +  
  theme_bw() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[0] ~ (eggs ~ m^-2 ~ day^-1))) + xlab("")+ 
  scale_colour_brewer(name = "Model",palette = "Dark2")+
  theme(legend.position = "none") +facet_wrap(~factor("A. Western Stratum"))+
  theme(strip.background = element_rect(fill = "white",colour = "black",strip.background = element_blank(), strip.text = element_text(face = "bold"))) 

Str2_resultsPo <- Str2_resultsPo %>% select(-dataset_short)


Str1_resultsZ$dataset_short <- paste0("20", sub("^Str1_", "", Str1_resultsZ$dataset))

plot_z_Str1<-ggplot() +
  geom_point(data = Str1_resultsZ,
             aes(y = estimate, x = dataset_short, col = model),
             size = 1, position=position_dodge(width = 0.65)) + 
  geom_errorbar(data = Str1_resultsZ, 
                aes(x = dataset_short, ymin = ci_lower, ymax = ci_upper, color = model), 
                width = 0.2,linewidth = 0.5, position=position_dodge(width = 0.65)) +  
  xlab("")+ ylab("")+
  theme_bw() +
 theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_colour_brewer(name = "Model",  palette = "Dark2")+
  theme(legend.position = "none")


Str1_resultsZ <- Str1_resultsZ %>% select(-dataset_short)

Str2_resultsZ$dataset_short <- paste0("20", sub("^Str2_", "", Str2_resultsZ$dataset))

plot_z_Str2<-ggplot() +
  geom_point(data = Str2_resultsZ,
             aes(y = estimate, x = dataset_short, col = model),
             size =1, position=position_dodge(width = 0.65)) + 
  geom_errorbar(data = Str2_resultsZ, 
                aes(x = dataset_short, ymin = ci_lower, ymax = ci_upper, color = model),
                width = 0.2,linewidth = 0.5,position=position_dodge(width = 0.65))  +  
   theme_bw() + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression (Z~(day^-1))) +xlab("")+
   scale_colour_brewer(name = "Model",palette = "Dark2")+
  theme(legend.position = "none")

Str2_resultsZ <- Str2_resultsZ %>% select(-dataset_short)

Str1_resultsPtot$dataset_short <- paste0("20", sub("^Str1_", "", Str1_resultsPtot$dataset))

plot_Ptot_Str1<- ggplot() +
  geom_point(data = Str1_resultsPtot, 
             aes(y = estimate/10^11, x = dataset_short, col = model),
             size=1,position=position_dodge(width = 0.65))+ 
  geom_errorbar(data = Str1_resultsPtot,
                aes(x= dataset_short, 
                    ymin = ci_lower /10^11, ymax = ci_upper/10^11, color = model), 
                  width = 0.2,linewidth = 0.5, position=position_dodge(width = 0.65)) +   
  ylab("")+xlab("")+  
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_colour_brewer(name = "Model",palette = "Dark2") +
  theme(legend.position = "none")

Str1_resultsPtot <- Str1_resultsPtot %>% select(-dataset_short)


Str2_resultsPtot$dataset_short <- paste0("20", sub("^Str2_", "", Str2_resultsPtot$dataset))

plot_Ptot_Str2<-ggplot() +
  geom_point(data = Str2_resultsPtot, 
             aes(y = estimate/10^11, x = dataset_short, col = model),
             size=1, position=position_dodge(width = 0.65)) + 
 geom_errorbar(data = Str2_resultsPtot, 
                aes(x= dataset_short, ymin = ci_lower /10^11, ymax = ci_upper/10^11, color = model), width = 0.2,linewidth = 0.5, position=position_dodge(width = 0.65)) +   
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(expression(P[tot] ~ (eggs ~ day^-1) ~x10^11 )) +xlab("")+
  scale_colour_brewer(name = "Model",  palette = "Dark2") +
  theme(legend.position = "none") 

Str2_resultsPtot <- Str2_resultsPtot %>% select(-dataset_short)

 
Summary_Plots<- 
  print(plot_grid(plot_Po_Str2,plot_Po_Str1,
                  plot_z_Str2, plot_z_Str1,
                  plot_Ptot_Str2, plot_Ptot_Str1,
                  ncol =  2))
         
model_legent_Str1 <- 
  get_legend(plot_Ptot_Str2 + 
               theme(legend.position = "bottom")+
               guides(color = guide_legend(nrow = 1))) 

final_Summary_Plots <- plot_grid(
   Summary_Plots,
   model_legent_Str1,
   ncol = 1,        
   rel_heights = c(1,0.05))
```

### DHAMR residuals 
```{r}
set.seed(123)
Residuals_test <- data.frame(
    Model = character(),
    Uniformity = character(),
    Uniformity_p = numeric(),
    Dispersion = character(),
    Dispersion_p= numeric(),
    Quantile   = character(),
    Quantile_p = numeric(),
    ZeroInfl   = character(),
    ZeroInfl_p = numeric(),
    SpatialAut = character(),
    SpatialAut_p= numeric())

resid_Chek <- function(model, data, a = 0.05, model_name = NULL) {
  
df_name <- deparse(substitute(data))
model_name <- paste0("model_M5", df_name)
  
# Clean data
data <- data %>% filter(!is.na(Abundance), !is.na(w_sn))

# Simulate and calculate residuals
s_fit <- simulate(model, nsim = 500, type = "mle-mvn")
r_fit <- dharma_residuals(s_fit, model, return_DHARMa = TRUE)

# Run DHARMa tests

res_tests<-DHARMa::testResiduals(r_fit,plot=T)  
q_test   <- DHARMa::testQuantiles(r_fit) # same as plotResiduals quantile residuals versus (ranked) fitted values. Check for homogeneity of variance-
zi_test  <- DHARMa::testZeroInflation(r_fit)

# For the spatial autocorelation 
simulationOutputS <- recalculateResiduals(
  r_fit,
  group = paste(data$X, data$Y),
  aggregateBy = mean)
  
sp_test <- DHARMa::testSpatialAutocorrelation(
    simulationOutputS,
    x = unique(data$X),
    y = unique(data$Y)  )
  
# Decide pass/fail
un_re  <- ifelse(res_tests$uniformity$p.value  < a, "Fail", "Pass")
dis_re  <- ifelse(res_tests$dispersion$p.value  < a, "Fail", "Pass")
q_res  <- ifelse(q_test$p.value  < a, "Fail", "Pass")
zi_res <- ifelse(zi_test$p.value < a, "Fail", "Pass")
sp_res <- ifelse(sp_test$p.value < a, "Fail", "Pass")
  
# Add results to global table
new_row <- data.frame(
  Model = paste(model_name),
  Uniformity = un_re,
  Uniformity_p = round(res_tests$uniformity$p.value,4),
  Dispersion = dis_re,
  Dispersion_p = round(res_tests$dispersion$p.value,4),
  Quantile = q_res,
  Quantile_p= round(q_test$p.value,4),
  ZeroInfl = zi_res,
  ZeroInfl_p= round(zi_test$p.value,4),
  SpatialAuto = sp_res,
  SpatialAuto_p= round(sp_test$p.value,4))
  
assign("Residuals_test", rbind(get("Residuals_test", envir = .GlobalEnv), new_row), envir = .GlobalEnv)

}

resid_Chek(model_M5Str1_06,Str1_06)
 #resid_Chek(model_M5Str2_06,Str2_06)

gt(Residuals_test,caption = "Residuals Test")
Residuals_test_M5 <- Residuals_test

#Do a  sum 
# Select only the test columns
test_cols <- c("Uniformity", "Dispersion", "Quantile", "ZeroInfl", "SpatialAuto")

# Function to count Pass
count_pass <- function(x) sum(x == "Pass")

# Create summary
summary_table <- data.frame(
  Test = test_cols,
  Pass = sapply(Residuals_test[test_cols], count_pass),
  Total = nrow(Residuals_test))

# Add Pass Rate as percentage
summary_table$PassRate <- round(summary_table$Pass / summary_table$Total * 100, 1)

# View the summary table
print(summary_table)
Residuals_test_M5_sum<-summary_table 
```

### Omegas
```{r}
all_omega_Str1 <- c(
  model_M5Str1_06_pred2$data$omega_s)

color_limits_Str1 <- c(min(all_omega_Str1, na.rm = TRUE),max(all_omega_Str1, na.rm = TRUE)) 

#all_omega_Str2 <- c(model_M5Str2_06_pred2$data$omega_s)
#color_limits_Str2 <- c(min(all_omega_Str2, na.rm = TRUE),max(all_omega_Str2, na.rm = TRUE)) 

#I do this with tiles as my surface is continuous.
tile_size <- 0.015
tile_width <- 1 / (111 * cos(25 * pi/180))  # ~0.0102°
tile_height <- 1 / 111  # ~0.0090°
tile_height <- tile_height * 1.15
tile_width <- tile_width * 1.2

plot_O_pub_Str1 <- function(prediction, m_code, legend = TRUE) {
  
p<- ggplot() +
   # Coastline
  geom_tile(
    data = prediction,
    aes(x = Lon, y = Lat, fill = omega_s),
    width = tile_width,  
    height = tile_height) +
    scale_fill_gradient2(name = expression(omega[s]),
     low = "red", mid = "white", high = "blue",
     midpoint = 0, 
     limits = color_limits_Str1)+
    labs(x = "Longitude", y = "Latitude", title = m_code) +
    theme(panel.background = element_rect(fill = "gray85", color = NA))+
    geom_sf(data = coast, fill = "gray20", col = "gray50") +
    coord_sf(xlim = c(23.5, 26), ylim = c(39.3, 41)) +
    scale_x_continuous(breaks = seq(23.5, 26.5, by = 1))
      if(!legend) {
   p <- p + theme(legend.position = "none")
  }
    print(p)
}
  
plot_O_pub_Str2 <- function(prediction, m_code, legend = TRUE) {
  
p<- ggplot() +
  geom_tile(
    data = prediction,
    aes(x = Lon, y = Lat, fill = omega_s),
    width = tile_width, 
    height = tile_height  ) +
    scale_fill_gradient2(name = expression(omega[s]),
     low = "red", mid = "white", high = "blue",
     midpoint = 0, 
     limits = color_limits_Str2)+
  labs(x = "Longitude", y = "Latitude", title = m_code) +
    theme(panel.background = element_rect(fill = "gray85", color = NA))+
    geom_sf(data = coast, fill = "gray20", col = "gray50") +
    coord_sf(xlim = c(22.5, 25), ylim = c(38.3, 40.9))+
    scale_x_continuous(breaks = seq(22.5, 25, by = 1))
if(!legend) {
   p <- p + theme(legend.position = "none")
  }
    print(p)
}

plotomega_Str1_06 <- plot_O_pub_Str1(model_M5Str1_06_pred2$data, "2006", legend = TRUE) 
#plotomega_Str2_06 <- plot_O_pub_Str2(model_M5Str2_06_pred2$data, "2006", legend = TRUE) 
```

### SVC_z
```{r}
#I do this with tiles as my surface is continious
tile_size <- 0.015
tile_width <- 1 / (111 * cos(25 * pi/180))  # ~0.0102°
tile_height <- 1 / 111  # ~0.0090°
tile_height <- tile_height * 1.15
tile_width <- tile_width * 1.2


plot_z_pub_Str1 <- function(prediction, m_code, legend = TRUE) {
    p<- ggplot() +
    geom_tile(
    data = prediction,
    aes(x = Lon, y = Lat, fill = -zeta_s_Age),
    width = tile_width,  
    height = tile_height  ) +
    scale_fill_gradient2(name = expression(z[s]),
     low = "red", mid = "white", high = "blue",
    midpoint = 0,
    guide = guide_colorbar( barwidth = unit(2.5, "mm"), barheight = unit(15, "mm")))+
    labs(x = "Longitude", y = "Latitude", title = m_code) +
    theme(panel.background = element_rect(fill = "gray85", color = NA))+
    geom_sf(data = coast, fill = "gray20", col = "gray50") +
    coord_sf(xlim = c(23.5, 26), ylim = c(39.3, 41)) +
   scale_x_continuous(breaks = seq(23.5, 26.5, by = 1))
  
     if(!legend) {
   p <- p + theme(legend.position = "none")
  }
    print(p)
}


plotz_06 <- plot_z_pub_Str1(model_M5Str1_06_pred2$data ,"2006", legend = TRUE)   

 
#I do this with tiles as my surface is continious
tile_size <- 0.015
tile_width <- 1 / (111 * cos(25 * pi/180))  # ~0.0102°
tile_height <- 1 / 111  # ~0.0090°
tile_height <- tile_height * 1.15
tile_width <- tile_width * 1.2


plot_z_pub_Str2 <- function(prediction, m_code, legend = TRUE) {
    p<- ggplot() +
    geom_tile(
    data = prediction,
    aes(x = Lon, y = Lat, fill = -zeta_s_Age),
    width = tile_width, 
    height = tile_height) +
    scale_fill_gradient2(name = expression(z[s]),
    low = "red", mid = "white", high = "blue",
    midpoint = 0,
    guide = guide_colorbar( barwidth = unit(2.5, "mm"), barheight = unit(15, "mm")))+
    labs(x = "Longitude", y = "Latitude", title = m_code) +
    theme(panel.background = element_rect(fill = "gray85", color = NA))+
    geom_sf(data = coast, fill = "gray20", col = "gray50") +
    coord_sf(xlim = c(22.5, 25), ylim = c(38.3, 40.9))+
    scale_x_continuous(breaks = seq(22.5, 25, by = 1))
  
     if(!legend) {
   p <- p + theme(legend.position = "none")
  }
    print(p)
}

#plotz_06 <- plot_z_pub_Str2(model_M5Str2_06_pred2$data ,"2006", legend = TRUE)   
```


# Overall comparisons 

### Fits plots
```{r}
Ploted_fits_Str1<- print(plot_grid(
    plot_M5_Str1_06 +
      labs (y= expression(P[ij]),x="", title= " ", subtitle="2006") + 
      scale_y_log10(labels = scales::label_number(),limits = c(1, NA)),ncol = 3))


Ploted_fits_Str2<- print(plot_grid(
plot_M4_Str2_06 +
  labs (y= expression(P[ij]),x="", title= " ", subtitle="2006") +
  scale_y_log10(labels = scales::label_number(),limits = c(1, NA)),
ncol = 3))
```


## Choose distibutions:
```{r}

data <- Str1_06 %>%filter(!is.na(Abundance), !is.na(w_sn))
pred_M3 <- predict(model_M3Str1_06, type = "response")
pred_M4 <- predict(model_M4Str1_06, type = "response")
pred_M5 <- predict(model_M5Str1_06, type = "response")
obs <- data$Abundance
rmse_Str1_06_M3 <- sqrt(mean((obs-pred_M3$est)^2, na.rm = TRUE))
rmse_Str1_06_M4 <- sqrt(mean((obs-pred_M4$est)^2, na.rm = TRUE))
rmse_Str1_06_M5 <- sqrt(mean((obs-pred_M5$est)^2, na.rm = TRUE))
cor_Str1_06_M3  <- cor(obs, pred_M3$est, use = "complete.obs", method = "pearson")
cor_Str1_06_M4  <- cor(obs, pred_M4$est, use = "complete.obs", method = "pearson")
cor_Str1_06_M5  <- cor(obs, pred_M5$est, use = "complete.obs", method = "pearson")


data <- Str2_06 %>%filter(!is.na(Abundance), !is.na(w_sn))
pred_M3 <- predict(model_M3Str2_06, type = "response")
pred_M4 <- predict(model_M4Str2_06, type = "response")
obs <- data$Abundance
  rmse_Str2_06_M3 <- sqrt(mean((obs-pred_M3$est)^2, na.rm = TRUE))
  rmse_Str2_06_M4 <- sqrt(mean((obs-pred_M4$est)^2, na.rm = TRUE))
  rmse_Str2_06_M5 <- NA
  rmse_Str2_06_M7 <- NA
  cor_Str2_06_M3  <- cor(obs, pred_M3$est, use = "complete.obs", method = "pearson")
  cor_Str2_06_M4  <- cor(obs, pred_M4$est, use = "complete.obs", method = "pearson")
  cor_Str2_06_M5 <- NA
```

## RMSE
```{r}
RMSE_sp<- tibble(dataset= c("Str1_06","Str2_06"),
        
        RMSE_M3= c(rmse_Str1_06_M3,rmse_Str2_06_M3),
    
        RMSE_M4= c(rmse_Str1_06_M4,rmse_Str2_06_M4),
    
        RMSE_M5= c(rmse_Str1_06_M5,rmse_Str2_06_M5),
        
        COR_M3= c(cor_Str1_06_M3, cor_Str2_06_M3),
        
        COR_M4= c(cor_Str1_06_M4,cor_Str2_06_M4),
        
        COR_M5= c(cor_Str1_06_M5,cor_Str2_06_M5),
        
        omega_M3 = c(omegaSD_M3_Str1_06,omegaSD_M3_Str2_06),
        
        omega_M4 = c(omegaSD_M4_Str1_06,omegaSD_M4_Str2_06),
        
        omega_M5 = c(omegaSD_M5_Str1_06,NA),
        
        zSD_M5 = c(svcSD_M5_Str1_06,NA),
        
        range_M3 = c(range_M3_Str1_06, range_M3_Str2_06),
        
        range_M4 = c(range_M4_Str1_06,range_M4_Str2_06),
        
        range_M5 = c(range_M5_Str1_06,NA),
        
        dispersion_M1 = c(dispersion_M1_Str1_06,dispersion_M1_Str2_06),
        
        dispersion_M2 = c(dispersion_M2_Str1_06,dispersion_M2_Str2_06),
        
        dispersion_M3 = c(dispersion_M3_Str1_06,dispersion_M3_Str2_06),
      
        dispersion_M4 = c(dispersion_M4_Str1_06,dispersion_M4_Str2_06),
        
        dispersion_M5 = c(dispersion_M5_Str1_06,NA))

```

```{r}
RMSE
RMSE_sp

df_rmse<-  tibble(dataset=RMSE_sp$dataset,
                 RMSE_M0=RMSE$RMSE_M0,RMSE_M1=RMSE$RMSE_M1, RMSE_M2=RMSE$RMSE_M2,
                 RMSE_M3=RMSE_sp$RMSE_M3,RMSE_M4=RMSE_sp$RMSE_M4,RMSE_M5=RMSE_sp$RMSE_M5,
                 Cor_M0= RMSE$COR_M0, Cor_M1= RMSE$COR_M1, Cor_M2= RMSE$COR_M2,
                 Cor_M3= RMSE_sp$COR_M3, Cor_M4= RMSE_sp$COR_M4,Cor_M5= RMSE_sp$COR_M5,
                 Disp_M1= RMSE_sp$dispersion_M1 , Disp_M2= RMSE_sp$dispersion_M2, Disp_M3= RMSE_sp$dispersion_M3, Disp_M4= RMSE_sp$dispersion_M4, Disp_M5= RMSE_sp$dispersion_M5)
             
df_rmse_long <- df_rmse %>%
  select(dataset,
       RMSE_M0, RMSE_M1, RMSE_M2, RMSE_M3, RMSE_M4,RMSE_M5) %>%pivot_longer(
    cols = -dataset,
    names_to = c(".value", "Model"),
    names_pattern = "(RMSE|Cor)_(.*)",
    values_to = "RMSE"
  )

df_rmse_long <- df_rmse_long %>%
  mutate(
    stratum = case_when(
      str_detect(dataset, "^Str1") ~ "B. Eastern Stratum",
      str_detect(dataset, "^Str2") ~ "A. Western Stratum"
    )  )

df_rmse_long <- df_rmse_long %>%
  mutate(
    year = as.integer(str_extract(dataset, "\\d{2}")),
    year = 2000 + year)

RMSE_plot<- 
ggplot(df_rmse_long,
       aes(x = as.factor(year), y = RMSE,
           group = Model,
           color = Model)) +
  geom_line(linewidth = 0.3,position=position_dodge(width = 0.25)) +
  geom_point(size = 1,position = position_dodge(width = 0.25)) +
scale_colour_brewer(
  name = "Model",
  palette = "Dark2")+
  facet_wrap(~stratum, ncol=1) +
  theme_bw() +
 theme(
    strip.background = element_rect(
    fill = "white",
    colour = "black",
  strip.background = element_blank(),
  strip.text = element_text(face = "bold"))) +
  theme(axis.text.x = element_text(angle = 40, hjust = 1)) +
  labs(
    x = "Year",
    y = "RMSE",
    color = "Model" ) 

RMSE_plot

```

```{r}
#Dispersion
df_disp_long <- df_rmse %>%
  select(dataset,
       Disp_M1, Disp_M2, Disp_M3, Disp_M4, Disp_M5) %>%pivot_longer(
    cols = -dataset,
    names_to = c(".value", "Model"),
    names_pattern = "(Disp)_(.*)",
    values_to = "Disp"  )

df_disp_long <- df_disp_long %>%
  mutate(
    stratum = case_when(
      str_detect(dataset, "^Str1") ~ "B. Eastern Stratum",
      str_detect(dataset, "^Str2") ~ "A. Western Stratum"   )  )

df_disp_long <- df_disp_long %>%
  mutate(
    year = as.integer(str_extract(dataset, "\\d{2}")),
    year = 2000 + year)

Disp_plot <- 
  ggplot(df_disp_long,
         aes(x = as.factor(year), y = Disp,
             group = Model,
             color = Model)) +
    # Background zones
    geom_rect(
        aes(xmin = -Inf, xmax = Inf, ymin = 0, ymax = 2, fill = "Negative Binomial"),
        alpha = 0.5,
        inherit.aes = FALSE     ) +
    geom_rect(
        aes(xmin = -Inf, xmax = Inf, ymin = 5.5, ymax = 15.5, fill = "Tweedie"),
        alpha = 0.5,
        inherit.aes = FALSE     ) +
  #  Fill legend 
  scale_fill_manual(
  name = "Distribution",
  values = c(
    "Negative Binomial" = scales::alpha("lightblue", 0.3),
    "Tweedie" = scales::alpha("lightpink", 0.3)))+
  # Lines and points
    geom_line(
        linewidth = 0.3,
        position = position_dodge(width = 0.25)) +
    geom_point(
        size = 1,
        position = position_dodge(width = 0.25)) +
  # Model colours
  scale_color_manual(
    name = "Model",
    values = RColorBrewer::brewer.pal(6, "Dark2")[2:6])+
  facet_wrap(~stratum) +
     # Primary and secondary axes
    scale_y_continuous(
        name = "φ",
        sec.axis = sec_axis(~ . / 10, name = "θ")) +
 theme_bw() +
 theme(
    strip.background = element_rect(
    fill = "white",
    colour = "black",
  strip.background = element_blank(),
  strip.text = element_text(face = "bold"))) +
theme(axis.text.x = element_text(angle = 45, hjust = 1),  axis.title.y = element_text(face = "italic")) +
    labs(
    x = "Year",
    color = "Model" )

Disp_plot
```


